### 2021-01-05 studies/2020_Aperol_Git/questionnaire data/gregor beispiele/Symlit_RMTS_pilot1_report/oldfiles_backup/Symlit_RMTS_plotting.Rmd
add examples

---
title: "Einzeldinge - Pilot and Preliminary Data"
author: "Gregor Kachel, Ramiro Glauer, Jan Lonnemann, Frauke Hildebrandt"
date: "03.04.2019"  
output: 
  bookdown::html_document2:
    toc: yes
    toc_float: true
    fig_caption: yes
    number_sections: false
    code_folding: hide
bibliography: library.bib
csl: apa.csl
header-includes:
  \usepackage{caption}
  \renewcommand{\thetable}{S\arabic{table}} 
  \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE)
```

```{r load_packages, include = FALSE}

# install.packages("tidyverse")
# install.packages("ggplot2")

library(ggplot2)
library(tidyverse)

```

```{r part 1-1-1 data, echo = F}

# load tab delimited text file into R; file should be in the same folder as the script for "./" to select it
raw.data<-read.table("./einzeltab.txt",sep="\t", header=T) 
# subsetting only data trials
raw.data<-raw.data[raw.data$content=="data",]
# subsetting only valid trials, exclude drops
raw.data<-raw.data[raw.data$valid=="in",]
# subsetting only first trial, exclude second trial
raw.data<-raw.data[raw.data$trial=="1",]
# subsetting only 12 to 28 months of age
raw.data<-raw.data[raw.data$agem<="30",]
# creating categories by splitting age in the middle
raw.data$splitage <- ifelse(raw.data$agem >= 18, 
                            c("18-27"), c("12-17")) 
# creating categories by making age-groups in the middle
attach(raw.data)
raw.data$agecat[agem <= 18] <- "12mo"
raw.data$agecat[agem > 18 & agem <= 23] <- "18mo"
raw.data$agecat[agem >= 24] <- "24mo"
detach(raw.data)
# factor to numeric
#raw.data$correct<-as.numeric(levels(raw.data$correct))[raw.data$correct] 
# drop levels and check exclusion
raw.data<-droplevels(raw.data)
# str(raw.data)

# creating data test with test, control1 and control2
test.data<-raw.data[raw.data$condition=="test",]
test.data<-droplevels(test.data)
control1.data<-raw.data[raw.data$condition=="control1",]
control1.data<-droplevels(control1.data)
control2.data<-raw.data[raw.data$condition=="control2",]
control2.data<-droplevels(control2.data)

# str(test.data)

```

### Abstract

We collected data from one- and two-year-olds (N = ; ages etc) in a set-up in which an object is established as being special by having the experimenter and child engage with it temporarily. Then, two additional objects are revealed that look just like the first one. All objects are successively placed in three locations in front of the child. While they are identical with regard to surface features, one of them is special due to the shared history the experimenter and child have with it. At test, the child is prompted to retrieve the special one from the set of three indistinguishable objects relying solely on its location. In order to show that one-year-olds can generally succeed in tracking an object in a set of three items, an additional control condition tests children with items that are distinguishable on the basis of surface features (here colour).

### Descriptive Analysis and State of Data Collection

Table with overview of data.
```{r part 1-1-2 data, echo = F}

# load tab delimited text file into R; file should be in the same folder as the script for "./" to select it
raw.data<-read.table("./einzeltab.txt",sep="\t", header=T) 
# subsetting only data trials
raw.data<-raw.data[raw.data$content=="data",]
# subsetting only valid trials, exclude drops
raw.data<-raw.data[raw.data$valid=="in",]
# subsetting only first trial, exclude second trial
 raw.data<-raw.data[raw.data$trial=="1",]
# subsetting only 12 to 28 months of age
raw.data<-raw.data[raw.data$agem<="30",]
# creating categories by splitting age in the middle
raw.data$splitage <- ifelse(raw.data$agem >= 18, 
                            c("18-27"), c("12-17"))

# creating categories by making age-groups in the middle
attach(raw.data)
raw.data$agecat[agem <= 17] <- "12mo"
raw.data$agecat[agem > 17 & agem <= 23] <- "18mo"
raw.data$agecat[agem <= 18] <- "12mo"
raw.data$agecat[agem > 18 & agem <= 23] <- "18mo"
raw.data$agecat[agem >= 24] <- "24mo"
detach(raw.data)

# factor to numeric
raw.data$correct<-as.numeric(levels(raw.data$correct))[raw.data$correct] 
# drop levels and check exclusion
raw.data<-droplevels(raw.data)
# str(raw.data)

overview<-aggregate(raw.data$correct, list(raw.data$agecat, raw.data$condition), mean)
N<-aggregate(raw.data$id, list(raw.data$agecat, raw.data$condition), length)
colnames(N)<- c("Age","Condition","lengthresult")
overview<- cbind(overview, N$lengthresult)
colnames(overview)<- c("Age","Condition","% correct", "N")
overview

# overview<-aggregate(raw.data$correct, list(raw.data$splitage, raw.data$condition), mean)
# N<-aggregate(raw.data$id, list(raw.data$splitage, raw.data$condition), length)
# colnames(N)<- c("Age","Condition","lengthresult")
# overview<- cbind(overview, N$lengthresult)
# colnames(overview)<- c("Age","Condition","% correct", "N")
# overview

```

### Test

```{r plot1, echo = FALSE, fig.cap="TEST CONDITION. Count of children's correct (green) and incorrect (red) choices from a set of three itmes plotted against their age in months. Line depicts smoothed conditional means. Error bars are 95% confidence intervals based on non-parametric bootstraps.One trial per participant.", fig.width = 10, fig.height = 5, fig.scale = 1.5}

#  create plot here 
test.data$correct.factor <- factor(test.data$correct)

g <- ggplot(test.data, aes(x=agem)) + 
  geom_dotplot(aes(fill=correct.factor), stackgroups = TRUE, binwidth = 0.5, binpositions = "all") + 
  geom_smooth(aes(y = correct, col=condition), method="lm") +
  scale_x_continuous(breaks=c(11:30))

g + labs(y="Proportion Correct", 
         x="Age in Months")

```

### Control 1

```{r part 1-1-3 data, echo = F}

# load tab delimited text file into R; file should be in the same folder as the script for "./" to select it
control1.data<-read.table("./einzeltab.txt",sep="\t", header=T) 
# subsetting only data trials
control1.data<-control1.data[control1.data$content=="data",]
# subsetting only valid trials, exclude drops
control1.data<-control1.data[control1.data$valid=="in",]
# subsetting only first trial, exclude second trial
control1.data<-control1.data[control1.data$trial=="1",]
# subsetting only 12 to 28 months of age
control1.data<-control1.data[control1.data$agem<="28",]
# subsetting only trials in the test condition
control1.data<-control1.data[control1.data$condition=="control1",]
# drop levels and check exclusion
control1.data<-droplevels(control1.data)
# creating categories by splitting age in the middle
control1.data$splitage <- ifelse(control1.data$agem >= 18, 
                            c("18-27"), c("12-17")) 
# creating categories by making age-groups in the middle
attach(control1.data)
control1.data$agecat[agem <= 17] <- "12mo"
control1.data$agecat[agem > 17 & agem < 24] <- "18mo"
control1.data$agecat[agem >= 24] <- "24mo"
detach(control1.data)

# factor to numeric
control1.data$correct<-as.numeric(levels(control1.data$correct))[control1.data$correct]

control1.data$correct.factor <- factor(control1.data$correct)

```

```{r plot2, echo = FALSE, fig.cap="CONTROL1 CONDITION. Count of children's correct (green) and incorrect (red) choices from a set of three itmes plotted against their age in months. Line depicts smoothed conditional means. Error bars are 95% confidence intervals based on non-parametric bootstraps.One trial per participant.", fig.width = 10, fig.height = 5, fig.scale = 1.5}
#  create plot here

g <- ggplot(control1.data, aes(x=agem)) + 
  geom_dotplot(aes(fill=correct.factor), stackgroups = TRUE, binwidth = 0.5, binpositions = "all") + 
  geom_smooth(aes(y = correct, col=condition), method="lm") +
  scale_x_continuous(breaks=c(11:30))
g + labs(y="Proportion Correct", 
         x="Age in Months")

```
### Control 2

```{r part 1-1-4 data, echo = F}

# load tab delimited text file into R; file should be in the same folder as the script for "./" to select it
control2.data<-read.table("./einzeltab.txt",sep="\t", header=T) 
# subsetting only data trials
control2.data<-control2.data[control2.data$content=="data",]
# subsetting only valid trials, exclude drops
control2.data<-control2.data[control2.data$valid=="in",]
# subsetting only first trial, exclude second trial
control2.data<-control2.data[control2.data$trial=="1",]
# subsetting only 12 to 28 months of age
control2.data<-control2.data[control2.data$agem<="28",]
# subsetting only trials in the test condition
control2.data<-control2.data[control2.data$condition=="control2",]
# drop levels and check exclusion
control2.data<-droplevels(control2.data)
# creating categories by splitting age in the middle
control2.data$splitage <- ifelse(control2.data$agem >= 18, 
                            c("18-27"), c("12-17")) 
# creating categories by making age-groups in the middle
attach(control2.data)
control2.data$agecat[agem <= 17] <- "12mo"
control2.data$agecat[agem > 17 & agem < 24] <- "18mo"
control2.data$agecat[agem >= 24] <- "24mo"
detach(control2.data)

# factor to numeric
control2.data$correct<-as.numeric(levels(control2.data$correct))[control2.data$correct]

control2.data$correct.factor <- factor(control2.data$correct)

```

```{r plot3, echo = FALSE, fig.cap="control2 CONDITION. Count of children's correct (green) and incorrect (red) choices from a set of three itmes plotted against their age in months. Line depicts smoothed conditional means. Error bars are 95% confidence intervals based on non-parametric bootstraps.One trial per participant.", fig.width = 10, fig.height = 5, fig.scale = 1.5}
#  create plot here

g <- ggplot(control2.data, aes(x=agem)) + 
  geom_dotplot(aes(fill=correct.factor), stackgroups = TRUE, binwidth = 0.5, binpositions = "all") + 
  geom_smooth(aes(y = correct, col=condition), method="lm") +
  scale_x_continuous(breaks=c(11:30))
g + labs(y="Proportion Correct", 
         x="Age in Months")

```