### 2021-01-22 studies/2020_Aperol_Git/questionnaire data/questionnaire_data_mandy.Rmd

---
title: "Pilot study APEROL: questionnaire data"
author: "Mandy Klatt, Gregor Kachel and Anne Deiglmayr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### HOW TO PREPARE DATA IN EXCEL ###
# use filter to select only lines with an ID in it 
# (i.e. exclude all empty lines or lines with na, ?, x, etc)
# select all columns except the columns for comments
# copy/paste into a new spreadsheet
# click "save as" and choose file type "tab-delimited-text"
####################################

# load tab delimited text file into R; 
# file should be in a subfolder named "data" for the script to select it
# import: never forget to insert "dec="," for replacing values with "," by "."
raw.data<-read.table("./data/short_questionnaire_data_2101211.txt", dec=",", sep="\t", header=T)

# loading packages
# install.packages("stringr") # install stringr to use the filter() function with text data
library(stringr)

# if a package is not installed on the current machine, it will install it
if (!require(tidyverse)) install.packages('tidyverse'); library(tidyverse)

# suppress "summarize" info. 
# if this line is committed, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

### checking for missing values
apply(is.na(raw.data), 2, which)

# knock out NAs
raw.data <- raw.data  %>% filter(
  personID != "NA")

# view (raw.data)
# str(raw.data)

```

# Short Summary
This report provides a summary of the questionnaire data of the first pilot APEROL. The aim was to explore connections between the visual attention of the teacher and students’ in-class experience of the teachers’ "withitness". The study was conducted as a lab study with simulated scripted mini-lessons (10min per lesson). Data was collected from N = 8 teachers in 2 sessions.

In each session, one subject took the role of teacher, while the other teachers acted as the class. After each lesson, the subjects answered items on classroom management, especially on the teacher's behaviour, using validated questionnaires. An external observer also answered a questionnaire. In addition, the teacher was asked to give a self-assessment on his/her classroom management.

### Data collection
For the pilot study, we collected questionnaire data from N = 8 subjects in `r sum(n_distinct(raw.data$session))` sessions. The questionnaires were answered after each lesson (10min) by an observer, the teacher and the three teachers acting as the class.

The table shows the state of data collection as of `r Sys.Date()`.

```{r overview demographics data table, echo=FALSE}

# years of experience
class(raw.data$experience)

# converting numbers to numeric
raw.data <- raw.data %>%  mutate(experience = as.numeric(as.character(factor(experience))))

# demographics and years of experience, grouped by group (N, gender, M age, max, min age)
raw.data %>%
  filter(what == "head") %>%
  group_by(group) %>%
  summarise(N=n(),
            "male" = sum(gender),
            "M age" = round(mean(age), 2),
            "min age" = min(age),
            "max age" = max(age),
            "SD age" = round(sd(age), 2),
            "M experience" = round(mean(experience), 2),
            "min experience" = min(experience),
            "max experience" = max(experience),
            "SD experience" = round(sd(experience), 2),
            ) %>% 
  knitr::kable(digits = 2)

```

### Plotting scales

The questionnaire contains following scales:

(1) classroom management
(2) positive climate and motivation
(3) clarity and structuredness
(4) activation and support
(5) presence: posture and gaze
(6) presence: voice
(7) presence: verbal and non-verbal intervention
(8) turn taking
(9) natural behaviour

```{r cleaning items, echo = FALSE}

# to compare both sessions filter only Identical items
raw.data <- raw.data  %>% filter(str_detect(shortID, 
                                                 paste(c("km_mitbekommen",
                                                         "km_klar",
                                                         "km_ungestört",
                                                         "km_aktiv",
                                                         "lkm_freundlich",
                                                         "lkm_interesse",
                                                         "lkm_kritik",
                                                         "lkm_ausreden",
                                                         "lkm_überlegen",
                                                         "lkm_rückmeldungen",
                                                         "ks_funktion",
                                                         "ks_verständlich",
                                                         "ks_sichtbar",
                                                         "af_beiträge",
                                                         "af_wechselseitig",
                                                         "af_nachdenken",
                                                         "phb_stand",
                                                         "phb_augen",
                                                         "phb_blick",
                                                         "phb_vorsichgeht",
                                                         "phb_raum",
                                                         "phb_gestik",
                                                         "phb_alleangesehen",
                                                         "ps_deutlich",
                                                         "ps_klar",
                                                         "ps_impulse",
                                                         "pi_nonverbal",
                                                         "pi_zubewegen",
                                                         "pi_direkt",
                                                         "redeanteil",
                                                         "m_natürlich",
                                                         "m_fiktiv",
                                                         "m_verhalten"),
                                                    collapse = '|')))

```

The scales on the quality of teaching are a validated questionnaire (Helmke 2014).Whereas the scales on the teacher's presence behaviour were derived from the research literature and were used in the pilot for the first time. 
For this reason, the various items for the scales ***presence*** are presented below.

### presence: posture and gaze

```{r mean presence: posture and gaze, echo = FALSE}

### presence: posture and gaze

# filter by parameter variable, i.e. create a subset for presence: posture and gaze
phb.data <- raw.data %>% filter(scale == "Präsenz/Haltung_Blick")

view (phb.data)

# value sometimes contained text before filtering, we have to convert the numbers to numeric
phb.data <- phb.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create  mean, min, max, sd of presence: posture/gaze
meanphb <- mean(phb.data$value)
minphb <- min(phb.data$value)
maxphb <- max(phb.data$value)
sdphb <- sd(phb.data$value)
Nphb <- phb.data %>% summarise(N = n_distinct(personID))

# filter by parametervariable, i.e. create a subset for experts and novice
expert.data <- phb.data %>% filter (group == "expert")
novice.data <- phb.data %>% filter (group == "novice")

# create mean, min, max, sd for expert/novice
# expert
expert.meanphb <- mean(expert.data$value)
expert.minphb <- min(expert.data$value)
expert.maxphb <- max(expert.data$value)
expert.sdphb <- sd(expert.data$value)
expert.Nphb<-length(unique(expert.data$personID))

#novice
novice.meanphb <- mean(novice.data$value)
novice.minphb <- min(novice.data$value)
novice.maxphb <- max(novice.data$value)
novice.sdphb <- sd(novice.data$value)
novice.Nphb<-length(unique(novice.data$personID))

```

The mean for all items of the scale ***presence: posture and gaze*** is `r round(meanphb, 2)` (min = `r minphb`; max = `r maxphb`; SD = `r round(sdphb,2)`; N = `r Nphb`; lickert-scale 1-4; 4 = best evaluation).

The mini-lessons of the experts were evaluated on average with `r round(expert.meanphb, 2)` (min = `r expert.minphb`; max = `r expert.maxphb`; SD = `r round(expert.sdphb,2)`; N = `r expert.Nphb`).

The mini-lessons of the novices were evaluated on average with `r round(novice.meanphb, 2)` (min = `r novice.minphb`; max = `r novice.maxphb`; SD = `r round(novice.sdphb,2)`; N = `r novice.Nphb`).

```{r presence: posture and gaze line plots,  fig.height = 4, message = FALSE, warning = FALSE , echo=FALSE}

phb.data$group <- as.factor(phb.data$group)

# to create error bars, we need to summarize the data in a separate data frame
phb.plot.sd <- phb.data %>%
  group_by(group, item.wordings, .drop=TRUE) %>%
  summarise(perspective, mean = mean(value), n = n(), sd = sd(value), se = sd/sqrt(n), )

names(phb.plot.sd)

# and here comes the long plot
phb.plot<-ggplot(data = phb.plot.sd,
                  aes(x = item.wordings, y = mean,
                      group = group, colour = group)) +
  geom_line()+
  geom_pointrange(position = position_dodge(0.1), aes(ymin = mean-se, ymax = mean+se))+
  theme_light() +
  ylim(1,4)+
  facet_grid(~perspective)+
  labs(data = phb.data, y = "value", x = NULL,
       title="presence: posture and gaze" ,
       subtitle=NULL)+
  theme(legend.position="bottom",
        panel.spacing.x = ,
        plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(labels = function(x) str_wrap(x, width = 55))+
  coord_flip()
phb.plot

```

### Overview all scales

```{r table all sclaes, echo=FALSE}

sum(is.na(raw.data$scale == "Aktivierung und Förderung"))

# value sometimes contained text before filtering, we have to convert the numbers to numeric
raw.data <- raw.data %>%  mutate(value = as.numeric(as.character(factor(value))))

# create a table of all scales 
raw.data %>% # select data
  group_by(group, scale) %>%
  summarise("M scale" = round(mean(value), 2),
            "min scale" = min(value),
            "max sclae" = max(value),
            "SD scale" = round(sd(value), 2),
            ) %>% 
  knitr::kable(digits = 2)

```

# References

Helmke, A., Helmke, T., Lenske, G., Pham, G., Praetorius, A. K., Schrader, F. W., & AdeThurow, M. (2014). Unterrichtsdiagnostik mit EMU. Aus-und Fortbildung der Lehrkräfte in Hinblick auf Verbesserung der Diagnosefähigkeit, Umgang mit Heterogenität und individuelle Förderung, 149-163.

R Core Team (2019). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686