### 2021-06-21 01_studies/02_2020_Aperol-pilot/questionnaire data/gregor beispiele/Symlit_RMTS_pilot1_report/Symlit_RMTS_pilot1_report.Rmd
SRI / Erhebungsprotokoll / Versuchsleitungsmanual

---
title: "Symlit RMTS Pilot1 Report"
author: "Gregor Kachel, Johanna Merker, Patrick Hawlitschek, Ivan Kroupin"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

### HOW TO PREPARE DATA IN EXCEL ###
# use filter to select only lines with an ID in it 
# (i.e. exclude all empty lines or lines with na, ?, x, etc)
# select all columns except the columns for comments
# copy/paste into a new spreadsheet
# click "save as" and choose file type "tab-delimited-text"
####################################

# load tab delimited text file into R; 
# file should be in a subfolder named "data" for the script to select it
raw.data<-read.table("./data/SymLit_RMTS_subjectlist_251220.txt",sep="\t", header=T)

# loading packages
# if a package is not installed on the current machine, it will install it
if (!require(tidyverse)) install.packages('tidyverse'); library(tidyverse)

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

# # change data structure: everything is turned into factors except for 
# main.data <- raw.data %>%
#   filter(what == "data"& valid == 1)  %>%
#   mutate(across(!correct&!agem&!aged, as.factor),
#          correct = as.integer(correct))

```

# Short Summary
This report provides a summary of the first pilot ran for SymLit_RMTS. Our aim for Pilot1 was to evaluate whether the tasks we are planning to use for SymLit_RMTS are appropriate for three- and four-year-old children. In a between-subjects design, three- and four-year-old children saw either a communicative framing of the task or a non-communicative task (condition (comm; non-comm) x agegroup (3yo; 4yo). Children were tested via online video calls. Since we were primarily interested in whether children understand the overall presentation of the task, we used very simple stimulus pairings in which the cue was a representation of the target stimulus (i.e. target is a circle; distractor is a square; cue is a drawn circle).

### Demographics
For pilot 1, we aimed at testing for four children in each condition and age-group. The table shows the state of data collection as of `r Sys.Date()`.

```{r overview demographics data table, echo=FALSE}
raw.data %>% # select data
  filter(what == "head") %>% # select only lines with data and only valid trials
  group_by(age, condition) %>%
  summarise(N = n_distinct(id),
            male = sum(sex),
            "mean age" = round(mean(agem), 2),
            "min age" = min(agem),
            "max age" = max(agem),
            "sd age" = round(sd(agem), 2),
            ) %>% 
  knitr::kable(digits = 2)

# demographics.data<- raw.data %>% # select data
#   filter(what == "head") %>% # select only lines with data and only valid trials
#   group_by(age, condition) %>%
#   summarise(N = n_distinct(id),
#             male = sum(sex),
#             "mean age" = round(mean(agem), 2),
#             "min age" = min(agem),
#             "max age" = max(agem),
#             "sd age" = round(sd(agem), 2),
#             )

```

### Performance in Familiarization Trials
All test sessions commence with a set of four familiarisation trials in which children have to point out the location of an item that was hidden in plain sight. These trials familiarize children with the task and ensure that the experimenter can read their responses (pointing) via the video call. Children need to succeed in at least half of the familiarization trials in order to progress to the test trials. In case children cannot point out the correct side of the screen, experimenters continue with a mock participation and participants would not be dropped from the sample.

```{r overview familiarization data table, echo=FALSE}
raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "fam") %>% 
  mutate(correct = as.integer(correct),
         valid = as.integer(valid))%>% 
  group_by(age, condition) %>%
  summarise(N = n_distinct(id),
            "trials/N" = round(sum(valid)/n_distinct(id), 2),
            "% correct" = round(mean(correct)*100, 2),
              ) %>% 
  knitr::kable(digits = 2)

```

### Performance in Test Trials
During the main phase of the study, children are presented with 16 test trials. In order to be submitted to the analyses, participants need to complete at least 8 test trials. Reasons for excluding trials include fuzziness and lack of attention, experimenter error, technical issues, and parental interference. Hence, the number of trials by participants can vary across conditions (cf. trials/N). Trials are coded as valid (valid = 1; drop = 0) and correct (correct = 1; error = 0). The table provides the percentages of correct choices during the main phase of the experiment.

```{r overview test data table, echo=FALSE}
raw.data %>% 
  filter(what == "data" & valid == "1" & trialtype == "test") %>% 
  mutate(correct = as.integer(correct),
         valid = as.integer(valid))%>% 
  group_by(age, condition) %>%
  summarise(N = n_distinct(id),
            "trials/N" = round(sum(valid)/n_distinct(id), 2),
            "% correct" = round(mean(correct)*100, 2),
              )  %>% 
  knitr::kable(digits = 2)

# test.data<-raw.data %>% # select data
#   filter(what == "data" & valid == "1" & trialtype == "test") %>% # select only lines with data and only valid trials
#   mutate(correct = as.integer(correct),
#          valid = as.integer(valid))%>% 
#   group_by(age, condition) %>%
#   summarise(N = n_distinct(id),
#             "trials/N" = round(sum(valid)/n_distinct(id), 2),
#             "% correct" = round(mean(correct)*100, 2),
#               )

# test.data<-raw.data %>% # select data
#   filter(what == "data" & valid == "1" & trialtype == "test") %>% # select only lines with data and only valid trials
#   mutate(correct = as.integer(correct),
#          valid = as.integer(valid))%>% 
#   group_by(age, condition) %>%
#   summarise(N = n_distinct(id),
#             "mean age" = round(mean(agem), 2),
#             "min age" = min(agem),
#             "max age" = max(agem),
#             #"sd age" = round(sd(agem), 2),
#             "trials/N" = round(sum(valid)/n_distinct(id), 2),
#             "% correct" = round(mean(correct)*100, 2)
#               ) 
# test.data %>% relocate("mean age", "min age", "max age", .after = "age")

```

### Proportion of Correct Choices
Graph provides boxplots and individual data for all experimental groups. Contrary to our hypothesis, children in both age-groups perform better in the non-communicative condition. The main reason behind this could be that the presentation of the communicative task is not clear enough. Mainly, we failed to get across that the monkey is actually drawing something for the children to help them. In the next phase we should make this more explicit. In general it needs to be noted that the number of participants per experimental group is very low. Other than that, data generally indicate that in both conditions and age-groups, children succeed in the task. Hence, our task in general and the video procedure in particular are age-appropriate and can be used to investigate our main research question.

```{r plot boxplot by groups, echo = FALSE}

test.ind.data <- raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "test") %>%
  select(id, age, condition, correct)  %>%
  mutate(correct = as.integer(correct),
         age = as.factor(age),
         condition = as.factor(condition),
         groups = paste(age,"-",condition)) %>%
  group_by(groups, age, condition, id) %>%
  summarise(proportion.correct = round(mean(correct)*100, 2))

test.all.data <- raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "test") %>%
  select(id, age, condition, correct)  %>%
  mutate(correct = as.integer(correct),
         age = as.factor(age),
         condition = as.factor(condition),
         groups = paste(age,"-",condition)) %>%
  group_by(groups, age, condition) %>%
  summarise(proportion.correct = round(mean(correct)*100, 2),
            sd = round(sd(correct)*100,2))

test.raw.data <- raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "test") %>%
  select(id, age, condition, correct)  %>%
  mutate(correct = as.integer(correct),
         age = as.factor(age),
         condition = as.factor(condition),
         groups = paste(age,"-",condition))

test.boxplot <-ggplot() +
  geom_boxplot(data = test.ind.data, aes(x = groups, y = proportion.correct), )+ # when there is more data add "notch = TRUE" for conf intervalls
  geom_jitter(data = test.ind.data, aes(x = groups, y = proportion.correct, colour=condition), size = 2, width = 0.1) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0,102))+
  labs(x="Groups", y="Proportion Correct")+ # for title add --> title="Proportion of Correct Choices", 
  theme_classic()
   # geom_violin(data = test.raw.data, trim = TRUE , adjust = .7, draw_quantiles = c(0.5), mapping = aes(x=groups, y = correct))
test.boxplot

```

### Performance across trials
Children are tested without feedback. The experimenter merely acknowledges the participant's choice by saying "thank you". The plot below presents the proportion of correct choices across all test trials for both age groups and conditions.

```{r plot linegraph performance across trials, echo=FALSE}

trial.data <- raw.data %>% 
  filter(what == "data" & valid == "1" & trialtype == "test") %>% 
  mutate(correct = as.integer(correct),
         valid = as.integer(valid),
         trial = as.integer(trial))%>% 
  select(age, condition, trial, correct) %>%
  group_by(age, condition, trial) %>%
    summarise(correct = round(mean(correct)*100, 2))

trial.plot <- ggplot(trial.data, aes(x = trial, y=correct, colour=condition, linetype=condition))+
  geom_line()+
  geom_point()+
  scale_x_continuous(breaks=trial.data$trial)+
  scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0,100))+
  labs(x="Trials", y="Proportion Correct")+ # to add a title --> title="Performance across test trials", 
  theme_classic() +
  facet_grid(age ~ .)
trial.plot

```

### Performance by items
In each trial, children were presented with a target and distractor shape as well as a cue matching the target shape. The counterbalancing ensures that all shapes occur both as target and distractor. Presentation in trials is counterbalanced. The graph below shows children's individual performance aggregated for all target shapes. More incorrect choices for a specific target could indicate that a cue is ambiguous.

```{r plot performance across items, echo=FALSE, warning=FALSE}

target.ind.data <- raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "test") %>%
  select(id, age, condition, tar, correct)  %>%
  mutate(correct = as.integer(correct),
         age = as.factor(age),
         condition = as.factor(condition),
         target = as.factor(tar),
         groups = paste(age,"-",condition)) %>%
  group_by(target,age, id) %>%
  summarise(proportion.correct = round(mean(correct)*100, 2))

target.boxplot <-ggplot() +
  geom_boxplot(data = target.ind.data, aes(x = target, y = proportion.correct),outlier.shape = NA)+ # when there is more data add "notch = TRUE" for conf intervalls
  geom_jitter(data = target.ind.data, aes(x = target, y = proportion.correct, colour=target), size = 2, width = 0.3) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0,110))+
  labs(x="Target Shapes", y="Proportion Correct")+ # for title add --> title="Proportion of Correct Choices", 
  theme_classic()+
  facet_grid(age ~ .)
   # geom_violin(data = test.raw.data, trim = TRUE , adjust = .7, draw_quantiles = c(0.5), mapping = aes(x=groups, y = correct))
target.boxplot

```

In addition, the plot below shows individual performance across all distractor shapes. More incorrect choices for a specific distractor might indicate that children prefer to select that distractor over any target.

```{r plot performance across distractor items, echo=FALSE, warning=FALSE}

distractor.ind.data <- raw.data %>%
  filter(what == "data" & valid == "1" & trialtype == "test") %>%
  select(id, age, condition, dis, correct)  %>%
  mutate(correct = as.integer(correct),
         age = as.factor(age),
         condition = as.factor(condition),
         distractor = as.factor(dis),
         groups = paste(age,"-",condition)) %>%
  group_by(distractor,age, id) %>%
  summarise(proportion.correct = round(mean(correct)*100, 2))

distractor.boxplot <-ggplot() +
  geom_boxplot(data = distractor.ind.data, aes(x = distractor, y = proportion.correct),outlier.shape = NA)+ # when there is more data add "notch = TRUE" for conf intervalls
  geom_jitter(data = distractor.ind.data, aes(x = distractor, y = proportion.correct, colour=distractor), size = 2, width = 0.3) +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 5), limits = c(0,110))+
  labs(x="Distractor Shapes", y="Proportion Correct")+ # for title add --> title="Proportion of Correct Choices", 
  theme_classic()+
  facet_grid(age ~ .)
   # geom_violin(data = test.raw.data, trim = TRUE , adjust = .7, draw_quantiles = c(0.5), mapping = aes(x=groups, y = correct))
distractor.boxplot

```

### Testing individual performance against chance
For each individual, we used a binomial test to investigate whether their performance was significantly above chance level (p < 0.05; 50%). The table provides the number and proportion of individuals performing above chance for age experimental group.

```{r table individual performance against chance, echo=FALSE}

ind.binom.data <- raw.data %>%
    filter(what == "data" & valid == "1" & trialtype == "test") %>% 
    select(condition, age, id, correct) %>% 
    mutate(correct = as.integer(correct))%>% 
    group_by(condition, age, id) %>%
    summarise("% correct" = round(mean(correct)*100, 2),
              "sum correct" = sum(correct),
              "sum trials" = length(correct),
              "p" = round(binom.test(sum(correct), length(correct), 
                               p = 0.5, alternative = "greater")$p.value, 3),
              "sig" = if_else(p < 0.05, 1, 0)
              )

sum.binom.data <- ind.binom.data %>%
    group_by(age, condition) %>%
    summarise(N = n_distinct(id),
            "succesfull"  = sum(sig))

sum.binom.data %>%
  mutate("% successful" = (succesfull/N)*100)%>% 
  knitr::kable(digits = 2)

```

### Testing group performance against chance
To test whether each children in both age-group and condition performed above chance on a group level, we used Wilcoxon / Mann-Whitney-U tests.

xxx To Do xxx

```{r plot performance against chance, echo=FALSE}

# Alternative using one-sample t-tests from emopoint
# this is not working yet as the participants in 4<o comm are all 100% correct
# raw.data %>%
#   filter(what == "data" & valid == "1" & trialtype == "test") %>% 
#   mutate(correct = as.integer(correct)) %>%
#   group_by(condition, age, id) %>%
#   arrange(condition, age) %>%
#   summarise(correct = mean(correct)) %>%
#   summarise(correct = list(correct)) %>%
#   group_by(condition, age) %>%
#   mutate(m = mean(unlist(correct)),
#          sd = sd(unlist(correct)),
#          df= t.test(unlist(correct), mu = 0.5)$parameter,
#          t_value = t.test(unlist(correct), mu = 0.5)$statistic,
#          p = t.test(unlist(correct), mu = 0.5)$p.value,
#          d = cohensD(unlist(correct), mu = 0.5))%>%
#   select(condition,age, m,sd,t_value,p,d) %>% #add df
#   knitr::kable(digits = 2)

```

### Comparing Performance across age-groups and conditions
We used Wilcoxon/Mann-Whitney-U-tests to test for significant differences in children's performance across age-groups and conditions respectively.

xxx To Do xxx

### Full Analysis (GLMM, ANOVA)
We are going to analyze our outcome (correct choice 0/1) in a 2x2 design (condition (comm, noncomm); age (3yo, 4yo) using either a GLMM or ANOVA depending on prior tests on data structure and model stability.

xxx To Do xxx

# References

R Core Team (2019). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.

Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686, https://doi.org/10.21105/joss.01686