### 2021-06-22 paper_1/plots_peer/presentation_plots.Rmd

---
title: "Plots"
author: 'Peer Schwertfeger'
date: |
  | peer.schwertfeger@uni-leipzig.de

  | 
  | \today
fontsize: 9 pt
output: 
  beamer_presentation:
    includes:
      in_header: preambel.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE,warning=FALSE,message=FALSE}
library(needs)
needs(tidyverse,
      gridExtra,
      viridis,
      ggthemes)

# prepare data (selected from questionnaire data)
df_gri <- read_delim("data/Aperol_pilot_glasses_raw_metrics_all_TOI.tsv", 
                     delim = "\t", escape_double = FALSE, trim_ws = TRUE)

# converting integer to numeric
df_gri <- mutate(.data = df_gri,
                 Duration_of_interval = as.numeric(Duration_of_interval),
                 Total_duration_of_whole_fixations = as.numeric(Total_duration_of_whole_fixations),
                 Number_of_whole_fixations = as.numeric(Number_of_whole_fixations),
                 Average_duration_of_whole_fixations = as.numeric(Average_duration_of_whole_fixations)
                 )

# calculate the GRI 
# create a basic table (tibble) using tidyverse functions
df_gri <- df_gri %>%
  group_by(TOI) %>%
  summarise("ID" = Participant,
            "Group" = Variable1,
            "DurTOI" = Duration_of_interval,
            "M Dur Fix" = mean(Total_duration_of_whole_fixations),
            "Number Fix" = Number_of_whole_fixations,
            "TimeTOI Dur" = Duration_of_interval/mean(Total_duration_of_whole_fixations),
            "OldGRI" = mean(Total_duration_of_whole_fixations)/Number_of_whole_fixations,
            "NewGRI" = (Duration_of_interval/mean(Total_duration_of_whole_fixations))/Number_of_whole_fixations,
            "NewGRI_scale" = scale((Duration_of_interval/mean(Total_duration_of_whole_fixations))/Number_of_whole_fixations))

df_gri$Group <- as_factor(df_gri$Group)

```

## Comparison of all variables

```{r}

df_gri_comp <- select(df_gri,
                      !ID)

df_gri_comp <- pivot_longer(data = df_gri,
                            c("DurTOI","M Dur Fix","Number Fix",
                              "TimeTOI Dur","OldGRI","NewGRI","NewGRI_scale"))

df_gri_comp$name <- factor(df_gri_comp$name,
                           levels = c("DurTOI","M Dur Fix","Number Fix",
                                      "TimeTOI Dur","OldGRI","NewGRI",
                                      "NewGRI_scale"),
                           labels = c("Duration TOI","Total dur. of whole fixations (mean)",
                                      "Number of whole fixations","TimeTOI Dur","Old GRI","Adjusted GRI","Adjusted GRI (scaled)")
                           )

ggplot(data = df_gri_comp,
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")

```

## Novice vs. Expert (Calibration)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Calibration phase"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```
## Novice vs. Expert (Asking Question)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Asking question"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Chatting with neighbour)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Chatting with neighbour"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Clicking pen)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Clicking pen"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Drawing on paper)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Drawing on paper"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Drumming hands)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Drumming hands"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Looking at phone)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Looking at phone"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Putting head on table)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Putting head on table"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (Yelling)

```{r message=FALSE,warning=FALSE}
ggplot(data = filter(df_gri_comp,
                     TOI == "Yelling"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  labs(y = "",
       x = ""
       ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name), 
             scale = "free")
```

## Novice vs. Expert (GRI)

```{r}
ggplot(data = filter(df_gri_comp,
                     name == "Old GRI" |
                     name == "Adjusted GRI" |
                     name == "Adjusted GRI (scaled)"),
       mapping = aes(x = Group,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  geom_point(size = 2, 
             alpha = 0.4,
             position = position_jitter(seed = 1, 
                                        width = 0.1
                                        )
             ) +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  facet_wrap(vars(name),
             scales = "free")
```

## Novice vs. Expert over TOIs

```{r}
ggplot(data = filter(df_gri_comp,
                     name == "Old GRI" |
                     name == "Adjusted GRI" |
                     name == "Adjusted GRI (scaled)"),
       mapping = aes(x = TOI,
                     y = value
                     )
       ) +
  geom_boxplot(mapping = aes(fill = Group)) +
  labs(x = "",
       y = "") +
  scale_fill_brewer(palette = "RdBu") +
  theme_clean() +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1
                                   )
        ) +
  facet_wrap(vars(name),
             scales = "free") 
```

## Descriptive Statistics

```{r message=FALSE,warning=FALSE,}

df_table <- tibble(Value = c("GRI","Adj. GRI","Adj. GRI (scale)"),
                   Mean = c(mean(df_gri$OldGRI),mean(df_gri$NewGRI),mean(df_gri$NewGRI_scale)),
                   SD = c(sd(df_gri$OldGRI),sd(df_gri$NewGRI),sd(df_gri$NewGRI_scale)),
                   Median = c(median(df_gri$OldGRI),median(df_gri$NewGRI),median(df_gri$NewGRI_scale))
                   )

test_data <- rnorm(n = length(df_gri$OldGRI),
                   mean = 0)

knitr::kable(df_table)

```