### 2021-07-21 02_paper/paper_1/plots_peer/preambel.tex
labellist

\usetheme[progressbar=frametitle,background=light,block=fill]{metropolis}
\usepackage{appendixnumberbeamer}

\setbeamertemplate{footline}
{
	\begin{beamercolorbox}[wd=0.5\textwidth,ht=3ex,dp=1.5ex,leftskip=.5em,rightskip=.5em]{author in head/foot}
		\usebeamerfont{author in head/foot}
		\hfill\insertshortauthor
	\end{beamercolorbox}
	\vspace*{-4.5ex}\hspace*{0.5\textwidth}
	\begin{beamercolorbox}[wd=0.5\textwidth,ht=3ex,dp=1.5ex,left,leftskip=.5em]{title in head/foot}
		\usebeamerfont{title in head/foot}
		\insertshorttitle
		\hfill \hfill \insertframenumber{} von \inserttotalframenumber
	\end{beamercolorbox}	
}

\setbeamertemplate{section in toc}[sections numbered]