### 2021-08-27 01_studies/01_Laborstudie ProVisioNET/R script/ProVisioNet_TOI-based_data.Rmd
Reliability test R Code

---
title: "TOI based ProVisioNet"
output: html_notebook
---

# Packages

```{r}
library(needs)
needs(tidyverse,
      psych)
```

# Read Data sets

```{r, include = FALSE}
df_EH <- read_tsv(file = "data/03_03_EH_ProVisioNET_study_glasses Metrics_interval-based.tsv"
                   )

df_RD <- read_tsv(file = "data/03_03_RD_ProVisioNET_study_glasses Metrics_interval-based.tsv"
                   )
```

# Data Wrangling

```{r}
df_EH <- df_EH %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

df_RD <- df_RD %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

# Start

df_cappa_start <- matrix(ncol = 3,
                         nrow = 8)

colnames(df_cappa_start) <- c("TOI","Start_of_interval_EH","Start_of_interval_RD")

df_cappa_start <- as_tibble(df_cappa_start)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_start$TOI <- TOI_names

# EH Start TOI
for (i in TOI_names) {

  val <- df_EH$Start_of_interval[df_EH == i]

  df_cappa_start$Start_of_interval_EH[df_cappa_start$TOI == i] <- val
}

# RD Start TOI
for (i in TOI_names) {

  val <- df_RD$Start_of_interval[df_RD == i]

  df_cappa_start$Start_of_interval_RD[df_cappa_start$TOI == i] <- val
}

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 8)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_EH","Duration_of_interval_RD")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# EH Duration TOI
for (i in TOI_names) {

  val <- df_EH$Duration_of_interval[df_EH == i]

  df_cappa_dur$Duration_of_interval_EH[df_cappa_dur$TOI == i] <- val
}

# RD Duration TOI
for (i in TOI_names) {

  val <- df_RD$Duration_of_interval[df_RD == i]

  df_cappa_dur$Duration_of_interval_RD[df_cappa_dur$TOI == i] <- val
}

```

# Reliabiltät

## Duration

```{r}
cor.test(x = df_cappa_dur$Duration_of_interval_EH[df_cappa_dur$TOI != "Lesson"],
     y = df_cappa_dur$Duration_of_interval_RD[df_cappa_dur$TOI != "Lesson"],
          method = "pearson")

ICC(x = select(df_cappa_dur,
               Duration_of_interval_EH,
               Duration_of_interval_RD
               )
    )

plot(x = df_cappa_dur$Duration_of_interval_EH[df_cappa_dur != "Lesson"],
     y = df_cappa_dur$Duration_of_interval_RD[df_cappa_dur != "Lesson"])
```

## Start

```{r}
cor.test(x = df_cappa_start$Start_of_interval_EH[df_cappa_start$TOI != "Lesson"],
     y = df_cappa_start$Start_of_interval_RD[df_cappa_start$TOI != "Lesson"],
          method = "pearson")

ICC(x = select(df_cappa_start,
               Start_of_interval_EH,
               Start_of_interval_RD
               )
    )

plot(x = df_cappa_start$Start_of_interval_EH[df_cappa_start != "Lesson"],
     y = df_cappa_start$Start_of_interval_RD[df_cappa_start != "Lesson"])
```