### 2021-09-07 01_studies/01_Laborstudie ProVisioNET/R script/ProVisioNet_ET_reli_01_01_TOI-based_data.Rmd
update R reaction video 01_01

---
title: "TOI based ProVisioNet"
output: html_notebook
---

# Packages

```{r}
if(!"devtools" %in% rownames(installed.packages())) install.packages("devtools")
if (!require(tidyverse)) install.packages('tidyverse'); library(tidyverse)
if (!require(papaja)) install.packages("papaja"); library(papaja)
if (!require(psych)) install.packages('psych'); library(psych) # stats
if (!require(moments)) install.packages('moments'); library(moments) # skewness & kurtosis
if (!require(sjPlot)) install.packages('sjPlot'); library(sjPlot) # item analysis of a scale or index
if (!require(DescTools)) install.packages('DescTools'); library(DescTools) # cohens kappa
if (!require(irr)) install.packages('irr'); library(irr) # Various Coefficients of Interrater Reliability and Agreement
if (!require(readxl)) install.packages('readxl'); library(readxl) # read in excel files

```

# Read Data sets

```{r, include = FALSE}
df_AP <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                   )

df_MK <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                   )
```

# Data Wrangling

```{r}
df_AP <- df_AP %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

df_MK <- df_MK %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

# Start

df_cappa_start <- matrix(ncol = 3,
                         nrow = 8)

colnames(df_cappa_start) <- c("TOI","Start_of_interval_EH","Start_of_interval_RD")

df_cappa_start <- as_tibble(df_cappa_start)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_start$TOI <- TOI_names

# AP Start TOI
for (i in TOI_names) {

  val <- df_AP$Start_of_interval[df_AP == i]

  df_cappa_start$Start_of_interval_AP[df_cappa_start$TOI == i] <- val
}

# MK Start TOI
for (i in TOI_names) {

  val <- df_MK$Start_of_interval[df_MK == i]

  df_cappa_start$Start_of_interval_MK[df_cappa_start$TOI == i] <- val
}

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 8)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_EH","Duration_of_interval_RD")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# AP Duration TOI
for (i in TOI_names) {

  val <- df_AP$Duration_of_interval[df_AP == i]

  df_cappa_dur$Duration_of_interval_AP[df_cappa_dur$TOI == i] <- val
}

# MK Duration TOI
for (i in TOI_names) {

  val <- df_MK$Duration_of_interval[df_MK == i]

  df_cappa_dur$Duration_of_interval_MK[df_cappa_dur$TOI == i] <- val
}

# In seconds

df_cappa_dur$Duration_of_interval_AP <- round(df_cappa_dur$Duration_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_dur$Duration_of_interval_MK <- round(df_cappa_dur$Duration_of_interval_MK/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_AP <- round(df_cappa_start$Start_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_MK <- round(df_cappa_start$Start_of_interval_MK/1000,
                                              digits = 2
                                             )
```

#
```{r}
fun_1 <- function(x) {
    y <- x
    return(y)
  }
```

# Reliabiltät

## Duration

```{r}
cor.test(x = df_cappa_dur$Duration_of_interval_AP,
     y = df_cappa_dur$Duration_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_dur,
               Duration_of_interval_AP,
               Duration_of_interval_MK
               )
    )

# Lesson is significant longer and gives a biased view on the points.

ggplot(data = filter(df_cappa_dur,
                     TOI != "Lesson"),
       mapping = aes(x = Duration_of_interval_AP,
                     y = Duration_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Duration coded by AP",
       y = "Duration coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```

## Start

```{r}
cor.test(x = df_cappa_start$Start_of_interval_AP,
     y = df_cappa_start$Start_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_start,
               Start_of_interval_AP,
               Start_of_interval_MK
               )
    )

ggplot(data = filter(df_cappa_start,
                     TOI != "Lesson"),
       mapping = aes(x = Start_of_interval_AP,
                     y = Start_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Start coded by AP",
       y = "Start coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```