### 2021-09-09 01_studies/01_Laborstudie ProVisioNET/R script/ProVisioNet_ET_EVENTS_01_01_TOI-based_data.Rmd
R script 02_02

---
title: "TOI based ProVisioNet"
output: html_notebook
---

# Packages

```{r}
library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl)
```

# Read Data sets

```{r, include = FALSE}
df_AP_01 <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_MK_01 <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_AP_02 <- read_tsv(file = "data/02_02_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_MK_02 <- read_tsv(file = "data/02_02_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")
```

# Data Wrangling

```{r}

df_AP_01 <- df_AP_01 %>% 
  select(TOI, Duration_of_interval)

df_MK_01 <- df_MK_01 %>% 
  select(TOI, Duration_of_interval)

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 11)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_AP_01","Duration_of_interval_MK_01")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase","Chatting_with_neighbour", "Clicking_pen","Drawing", "Drumming_with_hands",
               "Head_on_table", "Looking_at_phone", "Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# AP Duration TOI
for (i in TOI_names) {

  val <- df_AP_01$Duration_of_interval[df_AP_01 == i]

  df_cappa_dur$Duration_of_interval_AP_01[df_cappa_dur$TOI == i] <- val
}

# MK Duration TOI
for (i in TOI_names) {

  val <- df_MK_01$Duration_of_interval[df_MK_01 == i]

  df_cappa_dur$Duration_of_interval_MK_01[df_cappa_dur$TOI == i] <- val
}

# In seconds

df_cappa_dur$Duration_of_interval_AP_01 <- round(df_cappa_dur$Duration_of_interval_AP_01/1000,
                                              digits = 2
                                              )

df_cappa_dur$Duration_of_interval_MK_01 <- round(df_cappa_dur$Duration_of_interval_MK_01/1000,
                                              digits = 2
                                              )

```

#
```{r}
fun_1 <- function(x) {
    y <- x
    return(y)
  }
```

# Reliabiltät

## Duration

```{r}
cor.test(x = df_cappa_dur$Duration_of_interval_AP,
     y = df_cappa_dur$Duration_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_dur,
               Duration_of_interval_AP,
               Duration_of_interval_MK
               )
    )

# Lesson is significant longer and gives a biased view on the points.

ggplot(data = filter(df_cappa_dur,
                     TOI != "Lesson"),
       mapping = aes(x = Duration_of_interval_AP,
                     y = Duration_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_text(mapping = aes(label = TOI)
            ) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Duration coded by AP",
       y = "Duration coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```

## Start

```{r}
cor.test(x = df_cappa_start$Start_of_interval_AP,
     y = df_cappa_start$Start_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_start,
               Start_of_interval_AP,
               Start_of_interval_MK
               )
    )

ggplot(data = filter(df_cappa_start,
                     TOI != "Lesson"),
       mapping = aes(x = Start_of_interval_AP,
                     y = Start_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_text(mapping = aes(label = TOI)
            ) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Start coded by AP",
       y = "Start coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```