### 2021-09-10 01_studies/01_Laborstudie ProVisioNET/R script/ProVisioNet_ET_reli_01_01_04_04_AOI-based_data.Rmd
R script total duration AOIs

---
title: "AOI based ProVisioNet"
output: html_notebook
---

# Packages

```{r}
library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl)
```

# Read Data sets

```{r, include = FALSE}
AOI_AP_01 <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                   )

AOI_MK_01 <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                  )

AOI_AP_04 <- read_tsv(file = "data/04_04_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

AOI_MK_04 <- read_tsv(file = "data/04_04_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

```

# Data Wrangling

```{r}

# select relevant columns
AOI_AP_01 <- AOI_AP_01 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_MK_01 <- AOI_MK_01 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_AP_04 <- AOI_AP_04 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_MK_04 <- AOI_MK_04 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

# filter only for TOIs
AOI_AP_01 <- AOI_AP_01 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering" |
          TOI == "Entire Recording")

AOI_MK_01 <- AOI_MK_01 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering" |
          TOI == "Entire Recording")

AOI_AP_04 <- AOI_AP_04 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering"|
          TOI == "Entire Recording")

AOI_MK_04 <- AOI_MK_04 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering"|
          TOI == "Entire Recording")

# create new column names that give total duration of looking at all AOIs
AOI_AP_01 <- AOI_AP_01  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Material = Total_duration_of_fixations.Material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_MK_01 <- AOI_MK_01  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Teacher_Material = Total_duration_of_fixations.Material_Teacher,
          Students_Material = Total_duration_of_fixations.Material_Students,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_AP_04 <- AOI_AP_04  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Material = Total_duration_of_fixations.Material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_MK_04 <- AOI_MK_04  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Students_Material = Total_duration_of_fixations.Students_material,
          Teacher_Material = Total_duration_of_fixations.Teacher_material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

# replace NAs with 0
AOI_AP_01 <- AOI_AP_01 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_AP_04 <- AOI_AP_04 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_MK_01 <- AOI_MK_01 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_MK_04 <- AOI_MK_04 %>%
  mutate_all(~replace(., is.na(.), 0))

# create data set of total durations in long format for plotting --> for later
AOI_AP_01_long <- gather(AOI_AP_01, TOI, Duration)

AOI_MK_01_long <- gather(AOI_MK_01, TOI, Duration)

AOI_AP_04_long <- gather(AOI_AP_04, TOI, Duration)

AOI_MK_04_long <- gather(AOI_MK_04, TOI, Duration)

# create a new column with name sum_aois that gives the sum of looking at all AOIs
AOI_AP_01 <- AOI_AP_01  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

AOI_MK_01 <- AOI_MK_01  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Teacher_Material +
                                              Students_Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

################################################

# In seconds

df_cappa_dur$Duration_of_interval_AP <- round(df_cappa_dur$Duration_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_dur$Duration_of_interval_MK <- round(df_cappa_dur$Duration_of_interval_MK/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_AP <- round(df_cappa_start$Start_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_MK <- round(df_cappa_start$Start_of_interval_MK/1000,
                                              digits = 2
                                             )
```

#
```{r}
fun_1 <- function(x) {
    y <- x
    return(y)
  }
```

# Reliabiltät

## Duration

```{r}
cor.test(x = df_cappa_dur$Duration_of_interval_AP,
     y = df_cappa_dur$Duration_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_dur,
               Duration_of_interval_AP,
               Duration_of_interval_MK
               )
    )

# Lesson is significant longer and gives a biased view on the points.

ggplot(data = filter(df_cappa_dur,
                     TOI != "Lesson"),
       mapping = aes(x = Duration_of_interval_AP,
                     y = Duration_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_text(mapping = aes(label = TOI)
            ) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Duration coded by AP",
       y = "Duration coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```

## Start

```{r}
cor.test(x = df_cappa_start$Start_of_interval_AP,
     y = df_cappa_start$Start_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_start,
               Start_of_interval_AP,
               Start_of_interval_MK
               )
    )

ggplot(data = filter(df_cappa_start,
                     TOI != "Lesson"),
       mapping = aes(x = Start_of_interval_AP,
                     y = Start_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_text(mapping = aes(label = TOI)
            ) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Start coded by AP",
       y = "Start coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```