### 2021-09-13 01_studies/01_Laborstudie ProVisioNET/R script/ProVisioNet_ET_EVENTS_01_01_TOI-based_data.Rmd
R script TOI AOI

---
title: "TOI based ProVisioNet"
output: html_notebook
---

# Packages

```{r}
library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl)
```

# Read Data sets

```{r, include = FALSE}
df_AP_01 <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_MK_01 <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_AP_04 <- read_tsv(file = "data/04_04_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

df_MK_04 <- read_tsv(file = "data/04_04_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

```

# Data Wrangling

```{r}

df_AP_01 <- df_AP_01 %>% 
  select(TOI, Duration_of_interval)

df_MK_01 <- df_MK_01 %>% 
  select(TOI, Duration_of_interval)

# df_AP_02 <- df_AP_02 %>% 
#   select(TOI, Duration_of_interval)
# 
# df_MK_02 <- df_MK_02 %>% 
#   select(TOI, Duration_of_interval)

df_AP_04 <- df_AP_04 %>% 
  select(TOI, Duration_of_interval)

df_MK_04 <- df_MK_04 %>% 
  select(TOI, Duration_of_interval)

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 11)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_AP_01","Duration_of_interval_MK_01")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase","Chatting_with_neighbour", "Clicking_pen","Drawing", "Drumming_with_hands",
               "Head_on_table", "Looking_at_phone", "Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# AP_01 Duration TOI
for (i in TOI_names) {

  val <- df_AP_01$Duration_of_interval[df_AP_01 == i]

  df_cappa_dur$Duration_of_interval_AP_01[df_cappa_dur$TOI == i] <- val
}

# MK_01 Duration TOI
for (i in TOI_names) {

  val <- df_MK_01$Duration_of_interval[df_MK_01 == i]

  df_cappa_dur$Duration_of_interval_MK_01[df_cappa_dur$TOI == i] <- val
}

# AP_04 Duration TOI
for (i in TOI_names) {

  val <- df_AP_04$Duration_of_interval[df_AP_04 == i]

  df_cappa_dur$Duration_of_interval_AP_04[df_cappa_dur$TOI == i] <- val
}

# MK_04 Duration TOI
for (i in TOI_names) {

  val <- df_MK_04$Duration_of_interval[df_MK_04 == i]

  df_cappa_dur$Duration_of_interval_MK_04[df_cappa_dur$TOI == i] <- val
}

```

Erstellen von TOI_class
```{r}
df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Calibration_phase"] <- "Calibration_phase"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Chatting_with_neighbour" |
                         df_cappa_dur$TOI == "Heckling" |
                         df_cappa_dur$TOI == "Whispering"] <- "Verbal_disruption"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Head_on_table" |
                         df_cappa_dur$TOI == "Looking_at_phone" |
                         df_cappa_dur$TOI == "Drawing"] <- "Lack_of_eagerness"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Lesson"] <- "Lesson"

df_cappa_dur$TOI_class[is.na(df_cappa_dur$TOI_class)] <- "Agitation"

df_cappa_dur$TOI_class <- as_factor(df_cappa_dur$TOI_class)

df_cappa_dur_aggr <- df_cappa_dur %>%
  group_by(TOI_class) %>%
  summarise(Duration_mean_AP_01 = mean(Duration_of_interval_AP_01),
            Duration_mean_MK_01 = mean(Duration_of_interval_MK_01),
            Duration_mean_AP_04 = mean(Duration_of_interval_AP_04),
            Duration_mean_MK_04 = mean(Duration_of_interval_MK_04)
            )

# long format for plotting
dur_event_long  <- df_cappa_dur_aggr %>% gather(TOI, Duration)

# In seconds

df_cappa_dur_aggr$Duration_mean_AP_01 <- round(df_cappa_dur_aggr$Duration_mean_AP_01/1000,
                                              digits = 2
                                              )

df_cappa_dur_aggr$Duration_mean_MK_01 <- round(df_cappa_dur_aggr$Duration_mean_MK_01/1000,
                                              digits = 2
                                              )

df_cappa_dur_aggr$Duration_mean_AP_04 <- round(df_cappa_dur_aggr$Duration_mean_AP_04/1000,
                                              digits = 2
                                              )

df_cappa_dur_aggr$Duration_mean_MK_04 <- round(df_cappa_dur_aggr$Duration_mean_MK_04/1000,
                                              digits = 2
                                              )

```