### 2021-11-06 01_studies/01_Laborstudie ProVisioNET/Fitbit/FitBit Project/Pupil_dial.Rmd
pupil dial

---
title: "R Notebook"
output: html_notebook
---

# Packages

```{r}
library(needs)

needs(tidyverse,
      lubridate,
      hms,
      readxl)
```

# Data

```{r}
df_dia <- read_rds(file = "Data/df.rds")
```

```{r}
df_dia$date_time <- 
  make_datetime(hour = hour(df_dia$start_time),
                min = minute(df_dia$start_time),
                sec = second(df_dia$start_time)
                ) + 
  seconds(as.numeric(milliseconds(df_dia$time
                                          )
                             )
                )

df_dia$seconds <- floor(second(df_dia$date_time))
df_dia$minutes <- minute(df_dia$date_time)
df_dia$hours <- hour(df_dia$date_time)

df_dia_aggr <- 
df_dia %>% 
  group_by(hours, minutes, seconds, ID) %>%
  summarise(dia_left = mean(left),
            dia_right = mean(right)
            ) %>%
  ungroup() %>%
  mutate(Time = hms(hours = hours,
                    minutes = minutes,
                    seconds = seconds
                    )
         ) %>%
  select(dia_left,
         dia_right,
         Time,
         ID)

# df_dia_aggr$time <- 
# hms(hours = df_dia_aggr$hours,
#     minutes = df_dia_aggr$minutes,
#     seconds = df_dia_aggr$seconds)
#

rm(df_dia)

```

# Heart Rate data

```{r, message=FALSE, include=FALSE, error=FALSE, }
part <- read_xlsx(path = "Data/Heart Rate.xlsx")

#...............................................................................
#                                                                              .
#  Enter the missing cases in Heart Rate table below.                                       .
#                                                                              .
#...............................................................................

part <- filter(part,
              `ID participant` != 108 # |
              #`ID participant` != | etc.
              )

part <- part$`ID participant`

for (i in part) {
  nam <- paste("m", i, sep = "")
  assign(x = nam, 
         value = read_csv(file = paste("Data/", i, "_m.csv", 
                                       sep = ""
                                       )
                          )
         ) 
  }
```

# Extract data sets

```{r}
df_dia_aggr <-
  df_dia_aggr %>%
  filter(ID != "108" &
           ID != "109")

for (i in unique(df_dia_aggr$ID)) {

name <- paste0("pupil", i)
df <- filter(df_dia_aggr,
             ID == i)
assign(x = name,
       value = df)  
}
```

# concetanate data sets

```{r}

for (i in unique(df_dia_aggr$ID)) {

name1 <- paste0("m", i)
name2 <- paste0("pupil", i)
name3 <- paste0("tib_", i)
df <- left_join(x = get(name1),
                y = get(name2),
                by = "Time")

df <- df %>% mutate(dia_left = scale(dia_left),
                    dia_right = scale(dia_right),
                    `Heart Rate` = scale(`Heart Rate`)
                    )
assign(name3,
       value = df)  
}

rm(df)
rm(name3)

```

```{r}
df_compl <- 
  mget(ls(pattern = "tib")) %>%
  bind_rows() %>%
  mutate(ID = as_factor(ID),
         Type = as_factor(ifelse(test = as.numeric(as.character(ID)) > 200,
                                 yes = "advanced",
                                 no = "novice"
                                 )
                          )
         )
```

```{r}
ggplot(data = na.omit(df_compl),
       mapping = aes(x = dia_right,
                     y = `Heart Rate`
                     )
       ) +
  geom_point(mapping = aes(color = as_factor(Type))) +
  geom_smooth(method = "lm")

ggplot(data = na.omit(df_compl),
       mapping = aes(x = dia_left,
                     y = `Heart Rate`
                     )
       ) +
  geom_point(mapping = aes(color = as_factor(Type))) +
  geom_smooth(method = "lm")

ggplot(data = na.omit(df_compl)
       ) + 
  geom_histogram(mapping = aes(x = dia_left),
                 alpha = 0.4,
                 color = "blue") +
  geom_histogram(mapping = aes(x = dia_right),
                 alpha = 0.4,
                 color = "red")

```