### 2021-11-16 01_studies/01_Laborstudie ProVisioNET/R script/reli/ProVisioNet_ET_TOI_AOI-based_data.nb.Rmd
Coding data metrics + SRI + SJT

---
title: "AOI and TOI based analyses ProVisioNet"
output: html_notebook
---

# Install Packages

```{r}
library(needs)
needs(tidyverse,
      psych,
      moments,
      sjPlot,
      DescTools,
      irr,
      readxl,
      gridExtra)
```

# Read In Data sets

```{r, include = FALSE}
AOI_AP_01 <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                   )

AOI_MK_01 <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv"
                  )

AOI_AP_04 <- read_tsv(file = "data/04_04_AP_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

AOI_MK_04 <- read_tsv(file = "data/04_04_MK_ProVisioNET_study_glasses_Metrics_Intervall based.tsv")

```

# AOI based analyses 
# Data Wrangling - Total Duration and Proportional Duration of looking at AOIs

```{r, include = TRUE, cache = TRUE, echo=TRUE}

# select relevant columns
AOI_AP_01 <- AOI_AP_01 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_MK_01 <- AOI_MK_01 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_AP_04 <- AOI_AP_04 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

AOI_MK_04 <- AOI_MK_04 %>% 
  select(TOI,
         starts_with("Total_duration_of_fixation"))

# filter only for TOIs
AOI_AP_01 <- AOI_AP_01 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering" |
          TOI == "Entire Recording")

AOI_MK_01 <- AOI_MK_01 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering" |
          TOI == "Entire Recording")

AOI_AP_04 <- AOI_AP_04 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering"|
          TOI == "Entire Recording")

AOI_MK_04 <- AOI_MK_04 %>% 
  filter(  TOI == "Chatting_with_neighbour" |
           TOI == "Clicking_pen" |
           TOI == "Drawing" | 
           TOI == "Drumming_with_hands" |
           TOI == "Head_on_table" |
           TOI == "Looking_at_phone" |
           TOI == "Heckling"| 
           TOI == "Lesson" | 
           TOI == "Snipping_with_fingers" |
           TOI == "Whispering"|
          TOI == "Entire Recording")

# create new column names that give total duration of looking at all AOIs
AOI_AP_01 <- AOI_AP_01  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Material = Total_duration_of_fixations.Material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_MK_01 <- AOI_MK_01  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Teacher_Material = Total_duration_of_fixations.Material_Teacher,
          Students_Material = Total_duration_of_fixations.Material_Students,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_AP_04 <- AOI_AP_04  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Material = Total_duration_of_fixations.Material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

AOI_MK_04 <- AOI_MK_04  %>% 
  rename (Anna = Total_duration_of_fixations.Anna,
          Bianca = Total_duration_of_fixations.Bianca,
          Carl_a = `Total_duration_of_fixations.Carl(a)`,
          Disruptive_Person = Total_duration_of_fixations.Disruptive_Person,
          Board_Screen = Total_duration_of_fixations.Board_Screen,
          Classroom_Others = Total_duration_of_fixations.Classroom_Others,
          Students_Material = Total_duration_of_fixations.Students_material,
          Teacher_Material = Total_duration_of_fixations.Teacher_material,
          Nametag_Anna = Total_duration_of_fixations.Nametag_Anna,
          Nametag_Bianca = Total_duration_of_fixations.Nametag_Bianca,
          Nametage_Carl_a = `Total_duration_of_fixations.Nametag_Carl(a)`)

# replace NAs with 0
AOI_AP_01 <- AOI_AP_01 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_AP_04 <- AOI_AP_04 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_MK_01 <- AOI_MK_01 %>%
  mutate_all(~replace(., is.na(.), 0))

AOI_MK_04 <- AOI_MK_04 %>%
  mutate_all(~replace(., is.na(.), 0))

# create data set of total durations in long format for plotting --> for later
AOI_AP_01_long <- gather(AOI_AP_01, TOI, Duration)

AOI_MK_01_long <- gather(AOI_MK_01, TOI, Duration)

AOI_AP_04_long <- gather(AOI_AP_04, TOI, Duration)

AOI_MK_04_long <- gather(AOI_MK_04, TOI, Duration)

# create a new column with name sum_aois that gives the sum of looking at all AOIs
AOI_AP_01 <- AOI_AP_01  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

AOI_MK_01 <- AOI_MK_01  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Teacher_Material +
                                              Students_Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

AOI_AP_04 <- AOI_AP_04  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

AOI_MK_04 <- AOI_MK_04  %>% mutate(Sum_AOI = Anna + 
                                              Bianca +
                                              Carl_a + 
                                              Disruptive_Person + 
                                              Board_Screen + 
                                              Classroom_Others + 
                                              Teacher_Material +
                                              Students_Material + 
                                              Nametag_Anna +
                                              Nametag_Bianca + 
                                              Nametage_Carl_a)

# create new columns that give proportion of the duration of looking at all AOIs
# the proportion is the relative duration divided by total duration 
prop_AP_01 <- AOI_AP_01  %>% 
  mutate(Anna = Anna/Sum_AOI)   %>% 
  mutate(Bianca = Bianca/Sum_AOI)   %>% 
  mutate (Carl_a = Carl_a/Sum_AOI) %>%
  mutate(Disruptive_Person = Disruptive_Person/Sum_AOI)   %>% 
  mutate(Board_Screen = Board_Screen/Sum_AOI) %>%
  mutate(Classroom_Others = Classroom_Others/Sum_AOI) %>%
  mutate (Material = Material/Sum_AOI) %>%
  mutate (Nametag_Anna = Nametag_Anna/Sum_AOI) %>%
  mutate (Nametag_Bianca = Nametag_Bianca/Sum_AOI) %>%
  mutate (Nametage_Carl_a = Nametage_Carl_a/Sum_AOI)

# chose all columns exept the last one (Sum_AOI)
prop_AP_01 <- prop_AP_01 %>% filter(TOI == "Lesson") %>%
                              select(,1:11)

prop_MK_01 <- AOI_MK_01  %>% 
  mutate(Anna = Anna/Sum_AOI)   %>% 
  mutate(Bianca = Bianca/Sum_AOI)   %>% 
  mutate (Carl_a = Carl_a/Sum_AOI) %>%
  mutate(Disruptive_Person = Disruptive_Person/Sum_AOI)   %>% 
  mutate(Board_Screen = Board_Screen/Sum_AOI) %>%
  mutate(Classroom_Others = Classroom_Others/Sum_AOI) %>%
  mutate (Teacher_Material = Teacher_Material/Sum_AOI) %>%
  mutate (Nametag_Anna = Nametag_Anna/Sum_AOI) %>%
  mutate (Nametag_Bianca = Nametag_Bianca/Sum_AOI) %>%
  mutate (Nametage_Carl_a = Nametage_Carl_a/Sum_AOI) %>%
  rename(Material = Teacher_Material)

prop_MK_01 <- prop_MK_01 %>% filter(TOI == "Lesson") %>%
                              select(!"Students_Material")

prop_AP_04 <- AOI_AP_04  %>% 
  mutate(Anna = Anna/Sum_AOI)   %>% 
  mutate(Bianca = Bianca/Sum_AOI)   %>% 
  mutate (Carl_a = Carl_a/Sum_AOI) %>%
  mutate(Disruptive_Person = Disruptive_Person/Sum_AOI)   %>% 
  mutate(Board_Screen = Board_Screen/Sum_AOI) %>%
  mutate(Classroom_Others = Classroom_Others/Sum_AOI) %>%
  mutate (Material = Material/Sum_AOI) %>%
  mutate (Nametag_Anna = Nametag_Anna/Sum_AOI) %>%
  mutate (Nametag_Bianca = Nametag_Bianca/Sum_AOI) %>%
  mutate (Nametage_Carl_a = Nametage_Carl_a/Sum_AOI)

prop_AP_04 <- prop_AP_04 %>% filter(TOI == "Lesson") %>%
                              select(,1:11)

prop_MK_04 <- AOI_MK_04  %>% 
  mutate(Anna = Anna/Sum_AOI)   %>% 
  mutate(Bianca = Bianca/Sum_AOI)   %>% 
  mutate (Carl_a = Carl_a/Sum_AOI) %>%
  mutate(Disruptive_Person = Disruptive_Person/Sum_AOI)   %>% 
  mutate(Board_Screen = Board_Screen/Sum_AOI) %>%
  mutate(Classroom_Others = Classroom_Others/Sum_AOI) %>%
  mutate (Teacher_Material = Teacher_Material/Sum_AOI) %>%
  mutate(Students_Material = Students_Material/Sum_AOI) %>%
  mutate (Nametag_Anna = Nametag_Anna/Sum_AOI) %>%
  mutate (Nametag_Bianca = Nametag_Bianca/Sum_AOI) %>%
  mutate (Nametage_Carl_a = Nametage_Carl_a/Sum_AOI) %>%
  mutate (Material = Students_Material + Teacher_Material)

prop_MK_04 <- prop_MK_04 %>% filter(TOI == "Lesson") %>%
                              select(!"Students_Material") %>%
                              select(!"Teacher_Material") %>%
                              select (!"Sum_AOI")

# create data set of proportions in long format for plotting
prop_AP_01_long  <- prop_AP_01 %>% gather(TOI, Duration)

prop_MK_01_long <- prop_MK_01 %>% gather (TOI, Duration)

prop_AP_04_long <- prop_AP_04 %>% gather (TOI, Duration)

prop_MK_04_long <- prop_MK_04 %>% gather (TOI, Duration)

################################################

```

### Ploting data - Total Duration and Proportional Data of Video 01 and 04, coded by AP and MK

```{r echo = TRUE, include=TRUE}
total_AP_01 <- AOI_AP_01_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
            geom_jitter(color="black", size=0.4, alpha=0.9) +
            geom_point() + ylim (0,1000) + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Total Duration of Looking at AOIs in ms - Video 01 Rater AP") +
             xlab("AOI")

total_AP_01

total_MK_01 <- AOI_MK_01_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
            geom_jitter(color="black", size=0.4, alpha=0.9) +
            geom_point() + ylim (0,1000) + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Total Duration of Looking at AOIs in ms - Video 01 Rater MK") +
             xlab("AOI")

total_MK_01

### Plot data TOTAL DURATION video 04
total_AP_04 <- AOI_AP_04_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
            geom_jitter(color="black", size=0.4, alpha=0.9) +
            geom_point() + ylim (0,1000) + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Total Duration of Looking at AOIs in ms - Video 04 Rater AP") +
             xlab("AOI")

total_AP_04

total_MK_04 <- AOI_MK_04_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
            geom_jitter(color="black", size=0.4, alpha=0.9) +
            geom_point() + ylim (0,1000) + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Total Duration of Looking at AOIs in ms - Video 04 Rater MK") +
             xlab("AOI")

total_MK_04

### Plot data PROPORTIONAL DURATION video 01
prop_AP_01 <- prop_AP_01_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Proportional Duration of Looking at AOIs - Video 01 Rater AP") +
             xlab("AOI")

prop_AP_01

prop_MK_01 <- prop_MK_01_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Proportional Duration of Looking at AOIs in ms") +
             xlab("AOI")

prop_MK_01

### Plot data PROPORTIONAL DURATION video 04
prop_AP_04 <- prop_AP_04_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() +
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Proportional Duration of Looking at AOIs - Video 04 Rater AP") +
             xlab("AOI")

prop_AP_04

prop_MK_04 <- prop_MK_04_long %>%
            ggplot( aes(x=TOI, y=Duration, fill=TOI)) +
            geom_boxplot() + 
          theme(
            axis.text.x = element_text(angle=80, hjust=1),
            legend.position="none",
            plot.title = element_text(size=11)) +
          ggtitle("Proportional Duration of Looking at AOIs - Video 04 Rater MK") +
             xlab("AOI")

prop_MK_04

```

# Reliabilty for Hits in AOIs
# Read in Data Sets
```{r, include=TRUE}
AOI_hit_AP_01 <- read_tsv(file = "data/01_01_AP_ProVisioNET_study_glasses_Metrics_Event-based.tsv"
                   )

AOI_hit_MK_01 <- read_tsv(file = "data/01_01_MK_ProVisioNET_study_glasses_Metrics_Event-based.tsv"
                  )

AOI_hit_AP_04 <- read_tsv(file = "data/04_04_AP_ProVisioNET_study_glasses Metrics_event-based.tsv")

AOI_hit_MK_04 <- read_tsv(file = "data/04_04_MK_ProVisioNET_study_glasses Metrics_event-based.tsv")
```

# Data Wrangling
```{r echo=FALSE, include=TRUE}
# filter only rows lesson
AOI_hit_AP_01 <- AOI_hit_AP_01 %>% filter(TOI == "Lesson")

AOI_hit_MK_01 <- AOI_hit_MK_01 %>% filter(TOI == "Lesson")

AOI_hit_AP_04 <- AOI_hit_AP_04 %>% filter(TOI == "Lesson")

AOI_hit_MK_04 <- AOI_hit_MK_04 %>% filter(TOI == "Lesson")

# select variable AOI + Hit_proportion
AOI_hit_AP_01 <- AOI_hit_AP_01 %>% select(AOI, Hit_proportion)

AOI_hit_MK_01 <- AOI_hit_MK_01 %>% select(AOI, Hit_proportion)

AOI_hit_AP_04 <- AOI_hit_AP_04 %>% select(AOI, Hit_proportion)

AOI_hit_MK_04 <- AOI_hit_MK_04 %>% select(AOI, Hit_proportion)

# changing character to numeric
AOI_hit_AP_01$Hit_proportion <- as.numeric(AOI_hit_AP_01$Hit_proportion)

AOI_hit_MK_01$Hit_proportion <- as.numeric(AOI_hit_MK_01$Hit_proportion)

AOI_hit_AP_04$Hit_proportion <- as.numeric(AOI_hit_AP_04$Hit_proportion)

AOI_hit_MK_04$Hit_proportion <- as.numeric(AOI_hit_MK_04$Hit_proportion)

# round Hit_proportion to 1
AOI_hit_AP_01$Hit_proportion <- round(AOI_hit_AP_01$Hit_proportion, digits = 0)

AOI_hit_MK_01$Hit_proportion <- round(AOI_hit_MK_01$Hit_proportion, digits = 0)

AOI_hit_AP_04$Hit_proportion <- round(AOI_hit_AP_04$Hit_proportion, digits = 0)

AOI_hit_MK_04$Hit_proportion <- round(AOI_hit_MK_04$Hit_proportion, digits = 0)

#delete NAs
AOI_hit_AP_01 <- na.omit(AOI_hit_AP_01)

AOI_hit_MK_01 <- na.omit(AOI_hit_MK_01)

AOI_hit_AP_04 <- na.omit(AOI_hit_AP_04)

AOI_hit_MK_04 <- na.omit(AOI_hit_MK_04)

# group data by AOI and summarise Hit_prop
Hit_AP_01_aggr <- group_by(.data = AOI_hit_AP_01,
                    AOI) %>% summarise(Hit_count = length(Hit_proportion))

Hit_MK_01_aggr <- group_by(.data = AOI_hit_MK_01,
                    AOI) %>% summarise(Hit_count = length(Hit_proportion))

Hit_AP_04_aggr <- group_by(.data = AOI_hit_AP_04,
                    AOI) %>% summarise(Hit_count = length(Hit_proportion))

Hit_MK_04_aggr <- group_by(.data = AOI_hit_MK_04,
                    AOI) %>% summarise(Hit_count = length(Hit_proportion))

###### filter only for identical rows for both rater

AP_01_filter <- Hit_AP_01_aggr %>% filter(
                    AOI != "Nametag_Anna",
                    AOI != "Nametag_Bianca",
                    AOI != "Nametag_Carl(a)")

AP_01_filter

MK_01_filter <- Hit_MK_01_aggr %>% filter( 
                    AOI != "Bianca,Classroom_Others")

MK_01_filter$AOI[MK_01_filter$AOI == "Material_Teacher"] <- "Material"

MK_01_filter

# merge two data frames --> adding Columns, don't forget to merge BY AOI !

r3 <- merge(AP_01_filter, MK_01_filter, by = 'AOI')

# create a new df with only the ratings 
r3_agree <- r3 %>% select(Hit_count.x, Hit_count.y)

# function agree() with a tolerance
agree(r3_agree, tolerance=50)

#################### CohenKappa ##############################

# first, create a xtab and specify who is rater1 and rater2
ratertab <- xtabs(~r3$Hit_count.x + r3$Hit_count.y)
ratertab

# now you can calculate CohenKappa
CohenKappa(ratertab)

#################### ICC ##############################

# build a subset that contains the two AOI hit variables  
data_icc <- subset(r3, select = c(Hit_count.x, Hit_count.y))

# calculate the ICC for two raters with new subset
ICC(data_icc)

```

# TOI based analyses 
# Data Wrangling - TOIs classified by disruptions

```{r, echo=TRUE, include=TRUE}

df_AP_01 <- df_AP_01 %>% 
  select(TOI, Duration_of_interval)

df_MK_01 <- df_MK_01 %>% 
  select(TOI, Duration_of_interval)

df_AP_04 <- df_AP_04 %>% 
  select(TOI, Duration_of_interval)

df_MK_04 <- df_MK_04 %>% 
  select(TOI, Duration_of_interval)

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 11)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_AP_01","Duration_of_interval_MK_01")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase",
               "Chatting_with_neighbour",
               "Clicking_pen",
               "Drawing",
               "Drumming_with_hands",
               "Head_on_table",
               "Looking_at_phone",
               "Heckling","Lesson",
               "Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# AP_01 Duration TOI
for (i in TOI_names) {

  val <- df_AP_01$Duration_of_interval[df_AP_01 == i]

  df_cappa_dur$Duration_of_interval_AP_01[df_cappa_dur$TOI == i] <- val
}

# MK_01 Duration TOI
for (i in TOI_names) {

  val <- df_MK_01$Duration_of_interval[df_MK_01 == i]

  df_cappa_dur$Duration_of_interval_MK_01[df_cappa_dur$TOI == i] <- val
}

# AP_04 Duration TOI
for (i in TOI_names) {

  val <- df_AP_04$Duration_of_interval[df_AP_04 == i]

  df_cappa_dur$Duration_of_interval_AP_04[df_cappa_dur$TOI == i] <- val
}

# MK_04 Duration TOI
for (i in TOI_names) {

  val <- df_MK_04$Duration_of_interval[df_MK_04 == i]

  df_cappa_dur$Duration_of_interval_MK_04[df_cappa_dur$TOI == i] <- val
}

```

# Creating TOI classes
```{r, include=TRUE}
df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Calibration_phase"] <- "Calibration_phase"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Chatting_with_neighbour" |
                         df_cappa_dur$TOI == "Heckling" |
                         df_cappa_dur$TOI == "Whispering"] <- "Verbal_disruption"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Head_on_table" |
                         df_cappa_dur$TOI == "Looking_at_phone" |
                         df_cappa_dur$TOI == "Drawing"] <- "Lack_of_eagerness"

df_cappa_dur$TOI_class[df_cappa_dur$TOI == "Lesson"] <- "Lesson"

df_cappa_dur$TOI_class[is.na(df_cappa_dur$TOI_class)] <- "Agitation"

df_cappa_dur$TOI_class <- as_factor(df_cappa_dur$TOI_class)

df_cappa_dur_aggr <- df_cappa_dur %>%
  group_by(TOI_class) %>%
  summarise(Duration_mean_AP_01 = mean(Duration_of_interval_AP_01),
            Duration_mean_MK_01 = mean(Duration_of_interval_MK_01),
            Duration_mean_AP_04 = mean(Duration_of_interval_AP_04),
            Duration_mean_MK_04 = mean(Duration_of_interval_MK_04)
            )

```

#
```{r}
fun_1 <- function(x) {
    y <- x
    return(y)
  }
```

# TOI Reliability
# Start and Duration

# Data Wrangling

```{r, include=TRUE}
df_AP <- df_AP %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

df_MK <- df_MK %>% 
  select(TOI, Duration_of_interval, Start_of_interval)

# Start

df_cappa_start <- matrix(ncol = 3,
                         nrow = 8)

colnames(df_cappa_start) <- c("TOI","Start_of_interval_EH","Start_of_interval_RD")

df_cappa_start <- as_tibble(df_cappa_start)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_start$TOI <- TOI_names

# AP Start TOI
for (i in TOI_names) {

  val <- df_AP$Start_of_interval[df_AP == i]

  df_cappa_start$Start_of_interval_AP[df_cappa_start$TOI == i] <- val
}

# MK Start TOI
for (i in TOI_names) {

  val <- df_MK$Start_of_interval[df_MK == i]

  df_cappa_start$Start_of_interval_MK[df_cappa_start$TOI == i] <- val
}

# Duration

df_cappa_dur <- matrix(ncol = 3,
                       nrow = 8)

colnames(df_cappa_dur) <- c("TOI","Duration_of_interval_EH","Duration_of_interval_RD")

df_cappa_dur <- as_tibble(df_cappa_dur)

TOI_names <- c("Calibration_phase","Clicking_pen","Drumming_with_hands",
               "Head_on_table","Heckling","Lesson","Snipping_with_fingers",
               "Whispering")

df_cappa_dur$TOI <- TOI_names

# AP Duration TOI
for (i in TOI_names) {

  val <- df_AP$Duration_of_interval[df_AP == i]

  df_cappa_dur$Duration_of_interval_AP[df_cappa_dur$TOI == i] <- val
}

# MK Duration TOI
for (i in TOI_names) {

  val <- df_MK$Duration_of_interval[df_MK == i]

  df_cappa_dur$Duration_of_interval_MK[df_cappa_dur$TOI == i] <- val
}

# In seconds

df_cappa_dur$Duration_of_interval_AP <- round(df_cappa_dur$Duration_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_dur$Duration_of_interval_MK <- round(df_cappa_dur$Duration_of_interval_MK/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_AP <- round(df_cappa_start$Start_of_interval_AP/1000,
                                              digits = 2
                                              )

df_cappa_start$Start_of_interval_MK <- round(df_cappa_start$Start_of_interval_MK/1000,
                                              digits = 2
                                             )
```

```{r, include=FALSE}
fun_1 <- function(x) {
    y <- x
    return(y)
  }
```

# Reliabiltät

## Plotting Duration

```{r, echo=TRUE}
cor.test(x = df_cappa_dur$Duration_of_interval_AP,
     y = df_cappa_dur$Duration_of_interval_MK,
          method = "pearson")

ICC(x = select(df_cappa_dur,
               Duration_of_interval_AP,
               Duration_of_interval_MK
               )
    )

# Lesson is significant longer and gives a biased view on the points.

ggplot(data = filter(df_cappa_dur,
                     TOI != "Lesson"),
       mapping = aes(x = Duration_of_interval_AP,
                     y = Duration_of_interval_MK
                     )
       ) +
  geom_point(size = 2) +
  geom_text(mapping = aes(label = TOI)
            ) +
  geom_abline(intercept = 0,
              slope = 1,
              color = "red",
              linetype = 3
              ) +
  labs(x = "Duration coded by AP",
       y = "Duration coded by MK",
       title = "Inter-coder reliability"
       ) +
  theme_minimal()
```