### 2022-04-27 02_paper/02_study/paper_methods.Rmd
method paper new table

---
title             : "Through the eyes of the teacher - Multimodal exploration of expertise differences in the perception of classroom disruptions"
shorttitle        : "Professional Vision in Micro-Teaching-Units"

author: 
  - name          : "Mandy Klatt"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Dittrichring 5-7, 04109 Leipzig"
    email         : "mandy.klatt@uni-leipzig.de"
  - name          : "Dr. Gregor Kachel"
    affiliation   : "1, 2"
  - name          : "Dr. Christin Lotz"
    affiliation   : "1"
  - name          : "Prof. Dr. Anne Deiglmayr"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Leipzig University"
  - id            : "2"
    institution   : "Max-Planck University for Evolutionary Anthropology"

authornote: |

  We received funding from QualiFond of University Leipzig. We have no conflicts of interest to disclose. This article is based on data used at conference presentations (DACH-Nachwuchsakademie, 2022; EARLI SIG 11 2022; EARLI SIG 27, 2022).

abstract: |

keywords          : "Professional Vision, Expert-Novice-Paradigm, mobile Eye-Tracking"
wordcount         : "XXX"

bibliography      : ["r-references.bib"]

floatsintext      : no
figsintext        : yes
figurelist        : yes
tablelist         : yes
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
# bookdown::word_document2
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(include = FALSE)
library("papaja")
r_refs(file = "r-references.bib")
```

```{r analysis-preferences, include=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r include=FALSE}
# install.packages("needs")

# suppress "summarize" info. 
# if this line is ommitted, each table using the summarize function will be accompanied with a warning from the console
options(dplyr.summarise.inform = FALSE)

library(needs)

needs(tidyverse,
      lubridate,
      viridis,
      grid,
      gridExtra,
      cowplot,
      readxl,
      ARTofR,
      moments,
      ltm,
      sjPlot,
      kableExtra,
      xtable)

```

```{r data, include=FALSE}
# load data files

quest.data <- read_excel ("./data/data_empschul_labor_lehrperson.xlsx")

```

# Introduction

Managing classroom disruptions is a crucial aspect of effective classroom management (@evertson2006classroom; @kounin2006techniken).

Accordingly, teachers must be able to quickly notice and appropriately react to significant events in the classroom. This ability is referred to as classroom professional vision (@goodwin2015professional; @sherin2007development).

The process of professional vision can be divided into two main aspects: focusing on relevant situations for learning and teaching (“noticing”) and applying knowledge to draw appropriate conclusions in these situations (“knowledge-based reasoning”; @seidel2014modeling).

Therefore, the early visual perception of classroom disruptions is a key component to effectively maximize students’ learning time and minimize classroom interruptions. According to Kounin (2006), these important classroom management strategies are called “withitness” and “overlapping” and can be summarized under the concept of monitoring (@gold2017using).

Learning to develop such classroom management skills is a demanding and complex task for student teachers (@wolff2016teacher). Research on teacher expertise showed that expert and novice teachers differ in their ability to perceive classroom events, “[…] whereas only a few studies have focused on the basal process of noticing, i.e. the recognition of possible disturbing situations” (@grub2020process, p.75). Mobile eye-tracking data can fill this research gap by providing new insights in how expertise differences in teacher´s professional vision manifest in teacher-student interactions (@lachner2016makes; @@wolff2016teacher).

# Theoretical background

## Professional competence

## Classroom Management
- Disruptions defintiion

## Professional Vision

## Expertise

## Parameter/Indicators of professional vision

# Research Questions

This study examined how the degree of teaching experience influences (a) the number of fixations on relevant areas (e.g., the student performing the disruption), (b) the fixation duration in relevant areas and (c) the time to first fixation on relevant areas, using mobile eye-tracking data in a controlled, micro-teaching setting. Based on the existing literature, we expect expert teachers to outperform novices by (H1) showing more fixations on relevant areas with (H2) shorter fixation durations and (H3) perceiving classroom disruptions faster (cf. @van2014first).

# Methods
We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study. <!-- 21-word solution (Simmons, Nelson & Simonsohn, 2012; retrieved from http://ssrn.com/abstract=2160588) -->

## Participants

```{r demographicstable, include = TRUE, echo=FALSE}

# read in data and select and rename relevant columns with transmute 
demo.data <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% 
  transmute(Group = LI06_05,
                                 Gender = LI02_01_1,
                                 Age = LI03_01,
                                 `Teaching Experience` = LI04_01,
                                 `Type of School` = LI18,
                                 Grade = LI16_05,
                                 Subject = LI10_05,
                                 Semester = LI11_01,
                                 Internship = LI13_01,
                                 `Extracurricular Teaching Experience` = LI14,
                                 `Secondary Teaching Activities` = LI17
                                 )

# define expert and novice with ifelse function
demo.data$Group = ifelse(demo.data$Group < 200, "Novice","Expert")

# changing gender to count female gender (male = 2 --> 0)
demo.data$Gender = ifelse(demo.data$Gender == 2, "1","0")

# changing character Gender into numeric
demo.data$Gender <- as.numeric(demo.data$Gender)

# changing character Semester into numeric
demo.data$Semester <- as.numeric(demo.data$Semester)

# creating new column with percent of gender
demo.data <- demo.data %>%
  group_by(Group) %>%
  mutate(Gender_Sum=n())

# changing `Extracurricular Teaching Experience` to count experience (no experience = 2 --> 0)
demo.data$`Extracurricular Teaching Experience` = ifelse(demo.data$`Extracurricular Teaching Experience` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Extracurricular Teaching Experience` <- as.numeric(demo.data$`Extracurricular Teaching Experience`)

# changing `Secondary Teaching Activities` to count experience (no experience = 2 --> 0)
demo.data$`Secondary Teaching Activities` = ifelse(demo.data$`Secondary Teaching Activities` == 1, "1","0")

# changing character Extra. Exp. into numeric
demo.data$`Secondary Teaching Activities` <- as.numeric(demo.data$`Secondary Teaching Activities`)

# create new columns that give proportion of Gender
# percent = female gender divided by all female Gender
# demo.data <- demo.data  %>% 
#   group_by(Group) %>% 
#   mutate(Gender_percent = Gender/Gender_Sum) %>% 
#   mutate(Gender_percent = round(sum(Gender_percent)*100, digits=0))

# create a basic table (tibble) using tidyverse functions
demo.table <- demo.data %>%
  group_by(Group) %>%
  summarise(N=n(),
            "Gender female in percent" = round(sum(Gender/Gender_Sum) * 100, digits = 1),
            "M Age in years" = round(mean(Age), 1),
            "SD Age in years" = round(sd(Age), 1),
            "Min Age in years" = min(Age),
            "Max Age in years" = max(Age),
            "M Exp. in years" = round(mean(`Teaching Experience`), 1),
            "SD Exp. in years" = round(sd(`Teaching Experience`), 1),
            "Min Exp. in years" = min(`Teaching Experience`),
            "Max Exp. in years" = max(`Teaching Experience`),
            "M Semester" = round(mean(Semester), 1),
            "SD Semester" = round(sd(Semester), 1),
            "Min Semester" = min(Semester),
            "Max Semester" = max(Semester),
            "M Internship in teaching units" = round(mean(Internship), 1),
            "SD Internship in teaching units" = round(sd(Internship), 1),
            "Min Internship in teaching units" = min(Internship),
            "Max Internship in teaching units" = max(Internship),
            "Extracurricular Teaching Experience in percent" = round(sum(`Extracurricular Teaching Experience`/Gender_Sum) * 100, digits = 1),
            "Secondary Teaching Activities in percent" = round(sum(`Secondary Teaching Activities`/Gender_Sum) * 100)
           )

# format and insert table in manuscript
demo.table %>%
papaja::apa_table(
  caption = "Demographic Information and Teaching Experience",
  # note = "Write Note here",
  escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
  placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
  font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
  )

```

The sample consists of *N* = `r length(demo.data$Group)` participants with *n* = `r length(demo.data$Group[demo.data$Group == "Expert"])` expert teachers and *n* = `r length(demo.data$Group[demo.data$Group == "Novice"])` novice teachers.

The inclusion criterion for experts was that they have successfully completed teacher training and are actively employed in the teaching profession. According to @palmer2005identifying, we selected teachers as experts who had at least three years of professional experience and ideally had worked in another teaching position, such as subject advisor or trainer for trainee teachers, in addition to their teaching profession in school. Novices were student teachers who had successfully completed their first internship in a school and gained one to four hours of teaching experience.

The expert teachers (`r demo.data %>% filter(Gender == 1 & Group == "Expert") %>% pull(Gender) %>% length()` women; `r demo.table %>% filter(Group == "Expert") %>% pull("Gender female in percent")`%) had a mean age of `r demo.data %>% filter(Group == "Expert") %>% pull(Age) %>% mean() %>% round(., digits = 1)` years (*SD* = `r demo.table %>% filter(Group == "Expert") %>% pull("SD Age in years") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Expert") %>% pull("Min Age in years")`-`r demo.table %>%  filter(Group == "Expert") %>% pull("Max Age in years")`) and an average teaching experience of `r demo.table %>% filter(Group == "Expert") %>% pull("M Exp. in years") %>% mean() %>% round(., digits = 1)` years (*SD* = `r demo.table %>% filter(Group == "Expert") %>% pull("SD Exp. in years") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Expert") %>% pull("Min Exp. in years")`-`r demo.table %>%  filter(Group == "Expert") %>% pull("Max Exp. in years")`). `r demo.table %>% filter(Group == "Expert") %>% pull("Secondary Teaching Activities in percent")`% of the experienced teachers were also engaged in an secondary teaching activity, such as lecturers at the university, main training supervisors for trainee teachers and subject advisers.

The novice teachers (`r demo.data %>% filter(Gender == 1 & Group == "Novice") %>% pull(Gender) %>% length()` women; `r demo.table %>% filter(Group == "Novice") %>% pull("Gender female in percent")`%) had a mean age of `r demo.data %>% filter(Group == "Novice") %>% pull(Age) %>% mean() %>% round(., digits = 1)` years (*SD* = `r demo.table %>% filter(Group == "Novice") %>% pull("SD Age in years") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Novice") %>% pull("Min Age in years")`-`r demo.table %>%  filter(Group == "Novice") %>% pull("Max Age in years")`) with an average teaching experience of `r demo.table %>% filter(Group == "Novice") %>% pull("M Exp. in years") %>% mean() %>% round(., digits = 1)` years. On average, the student teachers were in their `r demo.table %>% filter(Group == "Novice") %>% pull("M Semester") %>% mean() %>% round(., digits = 1)` semester (*SD* = `r demo.table %>% filter(Group == "Novice") %>% pull("SD Semester") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Novice") %>% pull("Min Semester")`-`r demo.table %>%  filter(Group == "Novice") %>% pull("Max Semester")`). Furthermore, they had an average teaching experience of `r demo.table %>% filter(Group == "Novice") %>% pull("M Internship in teaching units") %>% mean() %>% round(., digits = 1)` teaching units à 45min (*SD* = `r demo.table %>% filter(Group == "Novice") %>% pull("SD Internship in teaching units") %>% round(., digits = 1)`; range: `r demo.table %>%  filter(Group == "Novice") %>% pull("Min Internship in teaching units")`-`r demo.table %>%  filter(Group == "Novice") %>% pull("Max Internship in teaching units")`) through the internships during their studies. `r demo.table %>% filter(Group == "Novice") %>% pull("Extracurricular Teaching Experience in percent")`% of the student teachers were also engaged in an extracurricular teaching activity, such as tutoring or homework supervision.

The subjects were primarily recruited through personal contacts, social media (Facebook), e-mail distribution lists and advertising in lectures at the University Leipzig. All study procedures were carried out in accordance with the ethical standards of the University’s Institutional Review Board. The authors received a positive vote on the study procedures from the Ethics Committee Board of Leipzig University. All participants were informed in detail about the aim and intention of the study prior to testing. Participation in the study was voluntary and only took place after written consent has been given.

## Material

### Eye-Tracking equipment

During the unit, teachers wore a binocular Tobii Pro Glasses 2 eye-tracker (https://www.tobiipro.com/product-listing/tobii-pro-glasses-2/). The system consisted of a wearable head unit and a recording unit. As shown in Figure \@ref(fig:tobiiglasses2), the head unit was a measuring device with different sensors. A high-definition scene camera captured a full HD video of the teacher’s field of vision. An integrated microphone recorded the surrounding sounds. Infrared light illuminators supported the eye tracking sensors which recorded the eye orientation to capure the tacher's gaze point as shown in Figure \@ref(fig:teachergazepoint). The videos were recorded with a sampling rate of 50 Hz in a video resolution with 1920 x 1080 at 25 frames per second. The scene camera had a field of view of 90 deg. in 16:9 format (82 deg. horizontal and 52 deg. vertical) and a frame dimension of 179 x 159 x 57 mm (width x depth x height).

```{r tobiiglasses2, include = T, fig.cap = "Tobii Pro Glasses 2; Source: https://www.tobiipro.com/product-listing/tobii-pro-glasses-2/", fig.align = "center", echo = FALSE}

knitr::include_graphics("./pictures/tobiiglasses2.png", dpi = 108)

```

```{r teachergazepoint, include = T, fig.cap = "Teacher's Gaze Point", fig.align = "center", echo = FALSE}

knitr::include_graphics("./pictures/teachergazepoint.png", dpi = 108)

```

###  Video and Audio recording equipment

The speech, sounds and voices of the participants were recorded with Zoom H3-VR Ambient Recorder (https://zoomcorp.com/en/gb/handheld-recorders/handheld-recorders/h3-vr-360-audio-recorder/) installed in the middle of the lab setting. The Zoom H3-VR recorded with four built-in mics arranged in an Ambisonic array with a bitrate of 4608 kBits/s.

Movements, facial expressions and gestures of the subjects were recorded by four Go Pro Hero 7 black cameras (https://gopro.com/content/dam/help/hero7-black/manuals/HERO7Black_UM_ENG_REVC.pdf) from different angles. The videos were recorded with a sampling rate of 50 Hz in a video resolution with 1920 x 1080 at 50 frames per second in 16:9 format with a linear field of view.

### Questionnaire

After each micro-teaching-unit, the students answered items on teaching quality using a validated questionnaire [@helmke2014unterrichtsdiagnostik] and scales on the teacher's presence behavior. In addition, participants were asked to give a self-assessment on  classroom management by completing the questionnaire after each micro-teaching-unit. The questionnaire was a 4-point Likert scale (1 = Strongly Disagree; 2 = Disagree; 3 = Agree; 4 = Strongly Agree).

```{r presence_questionnaire_novices, include=TRUE, echo=FALSE, results='asis'}
# load data files
quest.data <- read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% 
  mutate(Group = ifelse(LI06_05 < 200, "Novice","Expert")) %>% 
  dplyr::select(Group, LM01_01:LO01_02)

# create two data frames
quest.data.novice <- quest.data %>% filter(Group == "Novice")
quest.data.expert <- quest.data %>% filter(Group == "Expert")

############## NOVICE TABLE #########################

# creating a matrix
quest.table.novice <- matrix(data = NA, nrow = 4, ncol = 8)
rownames(quest.table.novice) <- c("LM", "LB", "LP", "LV")
  colnames(quest.table.novice) <- c("N Items","M","SD","Min","Max","Skewness","Kurtosis","alpha")

# creating a loop 
# i <- "LM"

for (i in c("LM", "LB", "LP", "LV")) {

loop.data <- quest.data.novice %>% 
  dplyr::select(starts_with(match = i))

quest.table.novice[i, "N Items"] <- length(loop.data)

# calculating mean
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "M"]

# calculating SD
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "SD"]

# calculating Min
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  min() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "Min"]

# calculating Max
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  max() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "Max"]

# calculating Skewness
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  skewness() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "Skewness"]

# calculating Kurtosis
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  kurtosis() %>% 
  round(., digits = 2) ->
quest.table.novice[i, "Kurtosis"]

# calculating alpha
alpha <- loop.data %>% 
  cronbach.alpha()

quest.table.novice[i, "alpha"] <- round(alpha$alpha,2)

}

# rename columns 
rownames(quest.table.novice) <- c("Classroom Management", "Balance", "Presence", "Natural Behavior")

# format and insert table in manuscript
papaja::apa_table(
quest.table.novice,
  caption = "Scale analysis for novices' self-assessment",
  # note = "Write Note here", 
  escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
  placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
  font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
)
```

```{r presence_questionnaire_experts, include=TRUE, echo=FALSE, results='asis'}

############## EXPERT TABLE #########################

# creating a matrix
quest.table.expert <- matrix(data = NA, nrow = 4, ncol = 8)
rownames(quest.table.expert) <- c("LM", "LB", "LP", "LV")
  colnames(quest.table.expert) <- c("N Items","M","SD","Min","Max","Skewness","Kurtosis","alpha")

# creating a loop 
# i <- "LM"

for (i in c("LM", "LB", "LP", "LV")) {

loop.data <- quest.data.expert %>% 
  dplyr::select(starts_with(match = i))

quest.table.expert[i, "N Items"] <- length(loop.data)

# calculating mean
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  mean() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "M"]

# calculating SD
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  sd() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "SD"]

# calculating Min
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  min() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "Min"]

# calculating Max
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  max() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "Max"]

# calculating Skewness
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  skewness() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "Skewness"]

# calculating Kurtosis
loop.data %>% 
  mutate(mean = rowMeans(.)) %>% 
  pull(mean) %>% 
  kurtosis() %>% 
  round(., digits = 2) ->
quest.table.expert[i, "Kurtosis"]

# calculating alpha
alpha <- loop.data %>% 
  cronbach.alpha()

quest.table.expert[i, "alpha"] <- round(alpha$alpha,2)

}

# rename columns 
rownames(quest.table.expert) <- c("Classroom Management", "Balance", "Presence", "Natural Behavior")

# format and insert table in manuscript
papaja::apa_table(
quest.table.expert,
  caption = "Scale analysis for experts' self-assessment",
  # note = "Write Note here", 
  escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
  placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
  font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
)

```

```{r item_analysis_classroom_management, include=TRUE, results='asis'}
quest.data %>% 
  dplyr::select(contains("LM")) %>% 
  as.data.frame() %>% 
  tab_itemscale(CSS = list(
    css.firsttablerow.col7 = "'cronbachs alpha';"
  ))
```

```{r item_analysis_balance, include=TRUE}
# quest.data %>% 
#   dplyr::select(contains("LB")) %>% 
#   as.data.frame() %>% 
#   tab_itemscale()
```

## Procedure

In June 2021, the study was piloted with student teachers volunteers to refine the study procedure. Data collection was conducted between July 2021, and July 2022.

## Data analysis

We investigated whether experts and novice teachers differed

All reported data analyses were conducted with the `r cite_r("r-references.bib")` and IBM SPSS 28.

# Results

# Discussion

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup